---
title: "APM_Cibersort"
author: "AJ"
date: "2025-08-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0. Install and load the required packages
```{r}
rm(list = ls())

library("tidyselect")
library("pillar")
library("data.table")
library("readr")
library("tibble")
library("AnnotationDbi")
library("org.Hs.eg.db")
library("pheatmap")
library("ggplot2")
library("openxlsx")
library("dplyr")
library("ggpubr")
library("patchwork")
library("ggrepel")
library("stringr")
library("tidyr")
library("textshape")
library("broom")
library("biomaRt")
library("scales")
library("ggsci")
library("broom")

PCAWG_path <- "/Users/wchen20/Desktop/PCAWG"
gritic_path <- paste0(PCAWG_path, "/MHC_evolution/GRITIC/output")
expr_path <- paste0(PCAWG_path, "/MHC_evolution/transcriptome/expression")
ciber_path <- paste0(PCAWG_path, "/MHC_evolution/transcriptome/Cibersort/abs")
matrix_path <- paste0(PCAWG_path, "/MHC_evolution/transcriptome/expression")

source("/Users/wchen20/Desktop/PCAWG/code/source/CIBERSORT.R")
source("/Users/wchen20/Desktop/PCAWG/code/source/EnrichedView_modified.R")
source("/Users/wchen20/Desktop/PCAWG/code/source/basic.R")
source("/Users/wchen20/Desktop/PCAWG/code/source/insert_newlines.R")
source("/Users/wchen20/Desktop/PCAWG/code/source/CRP_MAGEMSig0409.R")

ciber_cols <- c("B cells naive", "B cells memory", "Plasma cells", "T cells CD8", "T cells CD4 naive", 
                "T cells CD4 memory resting", "T cells CD4 memory activated", "T cells follicular helper", 
                "T cells regulatory (Tregs)", "T cells gamma delta", "NK cells resting", "NK cells activated", 
                "Monocytes", "Macrophages M0", "Macrophages M1", "Macrophages M2", "Dendritic cells resting", 
                "Dendritic cells activated", "Mast cells resting", "Mast cells activated", "Eosinophils", 
                "Neutrophils")

cibercom_cols <- c("T_cells_CD8", "T_cells_CD4", "B_cells", "NK_cells", "Macrophage", "Dendritic_cells", "Mast_cells", "Neutrophils", "Eosinophils")

LOH_MSI = LOH_mut <- fread(paste0(PCAWG_path, "/donors_and_biospecimens/LOH_MSI_final.csv")) 
LOH_type <- LOH_MSI %>% mutate(sample_id = aliquot_id) %>% dplyr::select(sample_id, histology_abbreviation)
LOH_aliquot <- LOH_MSI %>% mutate(sample_id = aliquot_id) %>% dplyr::select(sample_id, icgc_sample_id, MSI_mut)
sample_MSS <- LOH_MSI %>% filter(MSI_mut == "MSS") %>% pull(aliquot_id)

ciber_path <- "/Users/wchen20/Desktop/PCAWG/MHC_evolution/transcriptome/Cibersort/abs/250805/"
```

# 1. RUN CIBERSORT for TCGA RNA-seq data
## Combine all the gene expression for TCGA
```{r}
file_list <- list.files(path = matrix_path, pattern = "^exprSet_MAGE_", full.names = TRUE)

for (file in file_list) {
  tryCatch({
         count <- fread(file) %>% column_to_rownames(var = "V1") %>% as.matrix()
         file_name <- basename(file)
         type <- gsub("exprSet_MAGE_", "", file_name)
         type <- str_extract(type, "^[^.]+")

         sig_matrix <- system.file("extdata", "LM22.txt", package = "CIBERSORT")
         ciber_res <-  CIBERSORT(sig_matrix, count, perm = 0, QN = F, absolute = T) %>% as.data.frame()
         
         write.table(ciber_res, file = paste0(ciber_path, "/", type, "_ciber_abs.csv"), sep = ",", row.names = TRUE, quote = FALSE)
         
         
  }, error = function(e) {
  print(paste0("Error processing ", type, ": ", e$message))
  })
}
```

## Run CIBERSORT for single TCGA data
```{r}
ciber_path <- paste0(PCAWG_path, "/MHC_evolution/transcriptome/Cibersort/abs/241128/TCGA/single")
matrix_path <- "/Users/wchen20/Desktop/TCGA/RNAseq/matrix/"

sig_matrix_path <- "/Users/wchen20/Desktop/PCAWG/code/LM22.txt"
sig_matrix <- read.table(sig_matrix_path,header=T, sep="\t",row.names=1,check.names=F)

#The number of permutation
cibersort_perm = 100
#Quantile normalization of input mixture, default = FALSE for RNA-Seq data
cibersort_qn = FALSE
#whether to apply absolute mode in cibersort
cibersort_abs = TRUE
#sig.score = for each mixture sample, define S as the median expression,level of all genes in the signature matrix divided by the median expression level of all genes in the mixture. Multiple cell subset fractions by S.
cibersort_abs_method = "sig.score"

file_list <- list.files(path = matrix_path, pattern = "^exprSet_MAGE_", full.names = TRUE)

for (file in file_list) {
  
  tryCatch({
         count <- fread(file) %>% column_to_rownames("V1")
         file_name <- basename(file)
         type <- gsub("exprSet_MAGE_", "", file_name)
         type <- str_extract(type, "^[^.]+")
         
         ciber_res <- CIBERSORT(sig_matrix, 
                       count, 
                       perm = cibersort_perm, 
                       QN = cibersort_qn, 
                       absolute = cibersort_abs,
                       abs_method = cibersort_abs_method) %>% 
           as.data.frame()
         
         write.table(ciber_res, file = paste0(ciber_path, "/", type, "_ciber_abs.csv"), sep = ",", row.names = TRUE, quote = FALSE)
         
  }, error = function(e) {
  print(paste0("Error processing ", type, ": ", e$message))
  })
}
```

## Combine CIBCESORT results 
```{r}
file_list <- list.files(path = ciber_path, pattern = "_ciber_abs.csv", full.names = TRUE)
file_list <- grep("all", file_list, ignore.case = TRUE, value = TRUE, invert = TRUE)

ciber_list <- list()

for (file in file_list) {
  
  file_name <- basename(file)
  type <- gsub("_ciber_abs.csv", "", file_name)
  
  ciber <- fread(file) %>% dplyr::rename(sample_id = V1) %>% mutate(type = type)
  ciber_list[[type]] <- ciber
}

LOH_mut_gr <- LOH_MSI %>% dplyr::select(submitted_specimen_id, histology_abbreviation, aliquot_id)

ciber_all <- rbindlist(ciber_list) %>% 
  filter(substr(sample_id, 14, 15) %in% sprintf("%02d", 1:9)) %>%
  mutate(submitted_specimen_id = sub("^([^-]*-[^-]*-[^-]*-[^-]*).*", "\\1", sample_id)) %>% 
  inner_join(LOH_mut_gr, ., by = "submitted_specimen_id") %>%
  mutate(T_cells_CD8 = `T cells CD8`,
         T_cells_CD4 = `T cells CD4 naive`+`T cells CD4 memory resting`+`T cells CD4 memory activated`, 
         B_cells = `B cells naive` + `B cells memory`,
         NK_cells = `NK cells resting`+`NK cells activated`,
         Macrophage = `Macrophages M0` + `Macrophages M1` + `Macrophages M2`,
         Dendritic_cells = `Dendritic cells resting` + `Dendritic cells activated`,
         Mast_cells = `Mast cells resting` + `Mast cells activated`,
         Neutrophils = Neutrophils,
         Eosinophils = Eosinophils) %>%
  dplyr::select(-Correlation, -`P-value`, -RMSE, -type, -sample_id, -`Absolute score (sig.score)`) 
write.table(ciber_all, file = paste0(ciber_path, "/all_ciber_abs.csv"), sep = ",", row.names = FALSE, quote = FALSE)
```

## CIBCESORT results by APM events
```{r}
path_maf <- "/Users/wchen20/Desktop/PCAWG/consensus_snv_indel/maf/maf_from_vcf/"

mat1 <- fread(paste0(path_maf, "/onco_APM_hlaloh.txt")) %>% 
  column_to_rownames("V1") %>% 
  as.matrix()

long_df <- as.data.frame(mat1) %>%
  rownames_to_column("Gene") %>%
  pivot_longer(-Gene, names_to = "aliquot_id", values_to = "Status") %>%
  filter(Status != "") %>%         # only keep mutated entries
  mutate(APM_Event = "APM-Mutant") %>% 
  distinct(aliquot_id, .keep_all = T) %>% dplyr::select("aliquot_id", "APM_Event")
write.csv(long_df, file = "/Users/wchen20/Desktop/PCAWG/donors_and_biospecimens/APM_long_df.csv", row.names = F)

HLALOH_pcawg <- fread("/Users/wchen20/Desktop/PCAWG/donors_and_biospecimens/LOH_MSI_final.csv") %>% 
  dplyr::select(-APM_event) %>%
  subset(MSI_mut == "MSS") %>% 
  left_join(., long_df, by = "aliquot_id") %>% 
  mutate(APM_Event = ifelse(!is.na(APM_Event), "APM-Mutant", "APM-WT"),
         APM_Event = factor(APM_Event, levels = c("APM-WT", "APM-Mutant"))) %>% 
  dplyr::select("submitted_specimen_id", "histology_abbreviation", "APM_Event", "purity", "ploidy", "wgd_status", "total_perMB")
table(HLALOH_pcawg$APM_Event)
write.csv(HLALOH_pcawg, file = "/Users/wchen20/Desktop/PCAWG/donors_and_biospecimens/LOH_MSI_APM.csv", row.names = F)

ciber_all <- fread("/Users/wchen20/Desktop/PCAWG/MHC_evolution/transcriptome/Cibersort/abs/all_ciber_abs.csv") %>% 
  dplyr::select(-APM_Event) %>%
  inner_join(HLALOH_pcawg, by = c("submitted_specimen_id", "histology_abbreviation"))
write.csv(ciber_all, paste0(ciber_path, "/ciber_all.csv"), row.names = F)
```

## Compare the originial cell types
```{r}
ciber_all <- fread(paste0(ciber_path, "/ciber_all.csv"))

types <- unique(ciber_all$histology_abbreviation)

for (type in types) {
  
  ciber_type <- subset(ciber_all, histology_abbreviation == type) %>% dplyr::select(-submitted_specimen_id, -histology_abbreviation)
  
  data_long <- ciber_type %>% gather(Cell_type, Value_1, -APM_Event) %>% 
    mutate(Value_1 = as.numeric(Value_1),
           Cell_type = as.factor(Cell_type)) %>% 
    filter(!is.na(APM_Event))
  
  p_ciber <- ggplot(data_long, aes(x = Cell_type, y = Value_1, fill = APM_Event)) + 
    geom_violin(trim = FALSE, scale = "count", color = NA) +
    scale_fill_manual(values = my_colors) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 18, face = "bold", color = "black"),
          axis.title.y = element_text(size = 16),
          title = element_text(size = 20, face = "bold"),
          text = element_text(size = 16)) +
    labs(x = "", y = "Score", fill = "APM_event") +
    ggtitle(paste0("Immune Cell Abundance - ", type)) + 
    stat_compare_means(
      aes(group = APM_Event),
      method = "wilcox.test",
      label = "p.signif",
      size = 6)
  
  #ggsave(filename = paste0(ciber_path, "/", type, ".pdf"), plot = p_ciber, width = 30, height = 6.6, units = "in")
}

# All cancer types
# Filter the samples
filtered_data <- ciber_all %>%
  gather(key = "cell_type", value = "value", `B cells naive`:`Neutrophils`) %>%
  group_by(histology_abbreviation, cell_type) %>%
  filter(length(unique(APM_Event)) == 2)

# Ratio for median value
ratio_res <- filtered_data %>%
  group_by(histology_abbreviation, cell_type, APM_Event) %>%
  summarize(median_value = mean(value, na.rm = TRUE)) %>%
  spread(APM_Event, median_value) %>%
  mutate(ratio = `APM-Mutant` / `APM-WT`)

# Perform the Wilcoxon test
wilcox_res <- filtered_data %>%
  group_by(histology_abbreviation, cell_type) %>%
  do(tidy(wilcox.test(value ~ APM_Event, data = .))) 

# Perform linear regression
wilcox_res <- wilcox_res %>%
  group_by(histology_abbreviation) %>%
  mutate(fdr = p.adjust(p.value, method = "fdr")) 

wilcox <- left_join(ratio_res, wilcox_res, by = c("histology_abbreviation", "cell_type")) %>% 
          filter(!ratio == 0 & !is.infinite(ratio) & !is.nan(ratio)) %>%  
          group_by(histology_abbreviation) %>%
          mutate(fdr = p.adjust(p.value, method = "fdr"),
                 log10Ratio = log2(ratio),
                 `-log10FDR` = -log10(fdr)) %>%
          ungroup() 
                                                                                                    
significant_points <- subset(wilcox, fdr < 0.1)

plot_wilcox <- ggplot(wilcox, aes(x = histology_abbreviation, y = cell_type, color = log10Ratio, size = `-log10FDR`)) + 
  geom_point(alpha = 1) + 
  labs(x = "", y = "", color = "log10Ratio", size = "-log10FDR") +
  scale_color_gradient2(low = "#4DBBD5FF", high = "#DC0000FF", midpoint = 0, 
                        limit = c(-4, 4), 
                        breaks = c(-4, 0, 4),
                        labels = c("-4", "0", "4")) + 
  geom_text(data = significant_points, aes(label = "*"), vjust = 0.8, color = "black") +
  scale_size(range = c(0, 10)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 14, color = "black"),
        axis.text.y = element_text(size = 14, hjust = 1, color = "black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.background = element_blank())
ggsave(filename =  paste0(figure_path, "/Figure1d.pdf"), plot = plot_wilcox, width = 16, height = 7.5, units = "in")
```

## Combine cell types
```{r}
ciber_all <- fread(paste0(ciber_path, "/ciber_all.csv"))

filtered_data <- ciber_all  %>%
   mutate(T_cells_CD8 = `T cells CD8`,
         T_cells_CD4 = `T cells CD4 naive`+`T cells CD4 memory resting`+`T cells CD4 memory activated`, 
         B_cells = `B cells naive` + `B cells memory`,
         NK_cells = `NK cells resting`+`NK cells activated`,
         Macrophage = `Macrophages M0` + `Macrophages M1` + `Macrophages M2`,
         Dendritic_cells = `Dendritic cells resting` + `Dendritic cells activated`,
         Mast_cells = `Mast cells resting` + `Mast cells activated`,
         Neutrophils = Neutrophils,
         Eosinophils = Eosinophils) %>%
  dplyr::select(histology_abbreviation, APM_Event, all_of(cibercom_cols)) %>%
      pivot_longer(cols = all_of(cibercom_cols), 
               names_to = "cell_type", values_to = "value") %>%
  group_by(histology_abbreviation, cell_type) %>%
  filter(length(unique(APM_Event)) == 2)

# Ratio for median value
ratio_res <- filtered_data %>%
  group_by(histology_abbreviation, cell_type, APM_Event) %>%
  summarize(median_value = mean(value, na.rm = TRUE)) %>%
  spread(APM_Event, median_value) %>%
  mutate(ratio = `APM-Mutant` / `APM-WT`)

# Perform the Wilcoxon test
wilcox_res <- filtered_data %>%
  group_by(histology_abbreviation, cell_type) %>%
  do(tidy(wilcox.test(value ~ APM_Event, data = .))) 

wilcox_res <- wilcox_res %>%
  group_by(histology_abbreviation) %>%
  mutate(fdr = p.adjust(p.value, method = "fdr")) 

wilcox <- left_join(ratio_res, wilcox_res, by = c("histology_abbreviation", "cell_type")) %>% 
          filter(!ratio == 0 & !is.infinite(ratio) & !is.nan(ratio)) %>%  
          group_by(histology_abbreviation) %>%
          mutate(fdr = p.adjust(p.value, method = "fdr"),
                 log2Ratio = log2(ratio),
                 `-log10FDR` = -log10(fdr)) %>%
          ungroup() 
write.table(wilcox, file = paste0(ciber_path, "/ciber_wilcoxcom.csv"), sep = ",", row.names = FALSE, quote = FALSE)  
