---
title: "CRISPR_crossvalidation260220"
author: "AJ"
date: "2026-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0. Library packages
```{r}
library(openxlsx)

source("/Users/wchen20/Desktop/PCAWG/code/source/CRP_MAGEMSig260220.R")
```

# 1. MHC regulaters
## Paths
```{r}
path = "/Users/wchen20/Desktop/PCAWG/CRISPR/CRISPRscreen20240118/MHCregulators/"
path_save = "/Users/wchen20/Desktop/PCAWG/CRISPR/CRISPRscreen20240118/Crossvalidation/MHCregulators/"
```

## 2019_CancerCell_Burr -- 31564637
```{r, eval = FALSE}
# One side -- delete positive regulator
study <- "2019_CancerCell_Burr"
cells <- c("HLA", "HLA_B")
org = "hsa"
type = "MHC_high"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 4, type, i, path_save)
  
}
```

## 2021_Immunity_Dersh -- 33271120
```{r, eval = FALSE}
cells = c("BJAB", "HBL1", "SUDHL5", "TMD8")
study <- "2021_Immunity_Dersh"
org = "hsa"
type = "MHC"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

## 2023_Cell_Chen -- 37557169
```{r}
study <- "2023_Cell_Chen"
org = "hsa"
cells = c("HLA_A2_AFP", "HLA_ABC")
type = "MHC"


for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

## 2019_eLIFE_Pech -- 31452512
```{r, eval = FALSE}
study <- "2019_eLIFE_Pech"

cells <- "MHC"
org <- "hsa"
type <- "MHC"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

## 2019_CellReport_Freeman -- 31509742
```{r, eval = FALSE}
study <- "2019_CellReport_Freeman"

cells <- "MHClow"
org <- "mmu"
study <- "2019_CellReport_Freeman"
type = "MHC_low"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 4, type, i, path_save)
  
}
```

## 2021_CancerDiscovery_Gu -- 33589424
```{r}
study <- "2021_CancerDiscovery_Gu"
org = "mmu"
type = "MHC_low"
cells <- "MHC_low"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 3, type, i, path_save)
  
}

org = "mmu"
type = "MHC_high"
cells <- "MHC_high"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 3, type, i, path_save)
  
}
```

## 2023_NatureCellBiology_Sparbier 36635503
```{r, eval = FALSE}
study <- "2023_NatureCellBiology_Sparbier"

org = "hsa"
type = "MHC_low"
cells <- "HLA_Blow"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 4, type, i, path_save)
  
}

org = "hsa"
type = "MHC_high"
cells <- c("HLA_Bhigh", "HLA_ABChigh")

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 4, type, i, path_save)
  
}
```

# 2. Tcells
## Paths
```{r}
path = "/Users/wchen20/Desktop/PCAWG/CRISPR/CRISPRscreen20240118/Tcells/"
path_save = "/Users/wchen20/Desktop/PCAWG/CRISPR/CRISPRscreen20240118/Crossvalidation/Tcells/"
```

## 2019_Cell_Vredevoogd -- 31303383
```{r}
type = "Tcells"

study <- "2019_Cell_Vredevoogd"
cells <- c("IFNGR1deficient_D10_A", "IFNGR1deficient_D10_B")
org = "hsa"


for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

## 2020_Nature_Lawson -- 32968282
```{r}
study <- "2020_Nature_Lawson"
org = "mmu"
type ="Tcells"

cells <- getSheetNames(paste0(path,"/", study ,"/Paper/Table_S2_mTKO_normreadcounts.xlsx"))

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

## 2020_ScienceImmunology_Young -- 33443027
```{r, eval = FALSE}
study = "2020_ScienceImmunology_Young"
cells = "MC38"
org = "mmu"
type = "Tcells"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

## 2018_ScienceImmunology_Kearney -- 29776993
```{r, eval = FALSE}
study <- "2018_ScienceImmunology_Kearney"
cells <- c("B16_OTI", "B16_perforin_OTI", "MC38_IgG_OTI", "MC38_perforinko_OTI")
org = "mmu"
type = "Tcells"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 6, type, i, path_save)
  
}
```

## 2017_Nature_Patel -- 28783722
```{r, eval = FALSE}
study <- "2017_Nature_Patel"
cells <- c("Mel624")
org = "hsa"
type = "Tcells"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 6, type, i, path_save)
  
}
```

## 2023_CancerDiscovery_Ito -- 36811466
```{r, eval = FALSE}
study = "2023_CancerDiscovery_Ito"
cells = c("B16_OT1")
org = "mmu"
type = "Tcells"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

## 2018_Science_Pan -- 29301958
```{r, eval = FALSE}
study = "2018_Science_Pan"
cells = c("B16_OT_I", "B16_Pmel_1")
org = "mmu"
type = "Tcells"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 7, type, i, path_save)
  
}
```

## 2022_CellReportMedicine_Zhang -- 35688159
```{r, eval = FALSE}
study = "2022_CellReportMedicine_Zhang"
cells = c("D10_T")
org = "hsa"
type = "Tcells"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

## 2022_NatureCommunication_Frey -- 35379808
```{r, eval = FALSE}
study = "2022_NatureCommunication_Frey"
cells = c("KPC_OTI")
org = "mmu"
type = "Tcells"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

# 3. NK cell killing
## Paths
```{r}
path = "/Users/wchen20/Desktop/PCAWG/CRISPR/CRISPRscreen20240118/NKcells/"
path_save = "/Users/wchen20/Desktop/PCAWG/CRISPR/CRISPRscreen20240118/Crossvalidation/NKcells/"
```

## 2019_eLIFE_Pech -- 31452512
```{r, eval = FALSE}
cells = "K562"
org = "hsa"
study = "2019_eLIFE_Pech"
type = "NK"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

## 2021_NatureGenetics_Sheffer -- 34253920
```{r, eval = FALSE}
cells <- c("HT29", "SW620", "HCT15")
org <- "hsa"
study <- "2021_NatureGenetics_Sheffer"
type = "NK"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

## 2019_CellReport_Freeman -- 31509742
```{r, eval = FALSE}
### (control vs. treatment)
cells <- "B16_F10"
org = "mmu"
study = "2019_CellReport_Freeman"
type = "NK"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

## 2018_ScienceImmunology_Kearney -- 29776993
```{r, eval = FALSE}
cells <- c("MC38_NK")
org = "mmu"
study = "2018_ScienceImmunology_Kearney"
type = "NK"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 6, type, i, path_save)
  
}
```

## 2022_CellReportMedicine_Zhang -- 35688159
```{r, eval = FALSE}
study = "2022_CellReportMedicine_Zhang"
cells = c("D10_NK")
org = "hsa"
type = "NK"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

## 2024_CELLREPORT_Zhuang
```{r}
study = "2024_CELLREPORT_Zhuang"
cells = c("221")
org = "hsa"
type = "NKcells"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

## 2022_NC_Bernareggi
```{r}
study = "2022_NC_Bernareggi"
cells = c("GSC_387", "GSC_1517", "GSC_CW468", "GSC_D456")
org = "hsa"
type = "NKcells"
  
for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

## 2024_CELL_Li
```{r}
study = "2024_CELL_Li"
cells = c("COLO205_NK")
org = "hsa"
type = "NKcells"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

# 4. Macrophage killing
## Paths
```{r}
path = "/Users/wchen20/Desktop/PCAWG/CRISPR/CRISPRscreen20240118/Macrophages/"
path_save = "/Users/wchen20/Desktop/PCAWG/CRISPR/CRISPRscreen20240118/Crossvalidation/Macrophages/"
```

## 2021_Nature_Kamber -- 34497417
```{r, eval = FALSE}
cells <- "Ramos_macrophage"
org = "hsa"
study = "2021_Nature_Kamber"
type = "Macrophage"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```

## 2026_Science_Chung
```{r}
cells <- c("MOLM13_macrophage", "MV411_macrophage")
org = "hsa"
study = "2026_Science_Chung"
type = "Macrophage"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 8, type, i, path_save)
  
}
```

# 5. Gammadelta killing
## Paths
```{r}
path = "/Users/wchen20/Desktop/PCAWG/CRISPR/CRISPRscreen20240118/GammadeltaTcells/"
path_save = "/Users/wchen20/Desktop/PCAWG/CRISPR/CRISPRscreen20240118/Crossvalidation/GammadeltaTcells/"
```

## 2023_Nature_Mamedov -- 37648854 - negative should be checked
```{r, eval = FALSE}
cells <- "Daudi_Gammadelta"
org = "hsa"
study = "2023_Nature_Mamedov"
type = "Gammadelta"

for (i in c(50, 100, 200)) {
  
   CRP_MAGEMSig(cells, study, path, org, 5, type, i, path_save)
  
}
```