---
title: "2_3_Complex_Timing_of_Gains_Timelinedrivergene240806"
author: "AJ"
date: "2024-08-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0. Load packacges
```{r}
rm(list = ls())

library("stringr")
library("dplyr")
library("patchwork")
library("ggpubr")
library("ggridges")
library("truncnorm")
library("tidyr")
library("scatterpie")
library("tidyselect")
library("pillar")
library("data.table")
library("readr")
library("ggplot2")
library("openxlsx")
library("ggrepel")
library("tibble")
library("textshape")
library("matrixStats")

source(".../Project/code/source/basic.R")
source(".../Project/code/source/insert_newlines.R")
source(".../Project/code/source/plot_segTiming.R")

## Path
PCAWG_path <- "/rsrch6/scratch/hema_bio-Malignan/wchen20/PCAWG"
PCAWG_path <- ".../Project"
gritic_path <- paste0(PCAWG_path, "/MHC_evolution/GRITIC/output")

## Clinical data
LOH_MSI = LOH_mut <- fread(paste0(PCAWG_path, "/donors_and_biospecimens/LOH_MSI_final.csv"))

LOH_type <- LOH_MSI %>% mutate(sample_id = aliquot_id) %>% dplyr::select(sample_id, histology_abbreviation)
LOH_aliquot <- LOH_MSI %>% mutate(sample_id = aliquot_id) %>% dplyr::select(sample_id, icgc_sample_id, MSI_mut)
sample_MSS <- LOH_MSI %>% filter(MSI_mut == "MSS") %>% pull(aliquot_id)
type_list <- LOH_MSI %>% dplyr::select(aliquot_id, histology_abbreviation)

## Timing of segment gain
#seg <- fread(paste0(PCAWG_path, "/MHC_evolution/GRITIC/output/SNV_timing/wenjie_posterior_pcawg.tsv"))

## Timing of WGD
wgd_median <- fread(".../Project/MHC_evolution/GRITIC/output/wgd_median.csv") %>% dplyr::select(wgd_mediantiming, sample_id)

## Timing of non-synonymous mutations
nonsyn_filter_nonPloy <- fread(".../Project/MHC_evolution/GRITIC/output/filter/griticnon_filterclonal.tsv") %>%
  dplyr::select(-matches("\\.y$"), -Sample) %>%
  rename_with(~ str_replace_all(., "\\.x$", "")) %>% 
  filter(sample_id %in% sample_MSS)

sample_sucess <- unique(nonsyn_filter_nonPloy$sample_id)

## Save Path
gritic_path = ".../Project/MHC_evolution/GRITIC/output/250717/summary250717/"
path_timeline = ".../Project/MHC_evolution/GRITIC/output/250717/summary250717/timeline/"
```

# Update Polysover Timing
## Sample with HLA mutations
```{r}
maf_tcga <- fread(".../Project/MHC_evolution/HLAmutations/results/maf_tcga_HLAmut_final.maf")
maf_icgc <- fread(".../Project/MHC_evolution/HLAmutations/results/maf_icgc_HLAmut_final.maf")

## 71 HLA mutations in 61 samples
HLAall <- rbind(maf_tcga, maf_icgc, fill=TRUE) 
sample_HLA <- HLAall %>% pull(Sample) %>% unique()

## 57 SNVs in 61 samples
sample_HLAsnv <- HLAall %>% filter(!Variant_Classification %in% c("Frame_Shift_Del", "Frame_Shift_Ins", "In_Frame_Del", "In_Frame_Ins")) %>% 
  pull(Sample) %>% unique()
length(sample_HLAsnv)

## 5 Indels in 61 samples
sample_HLAindel <- HLAall %>% filter(Variant_Classification %in% c("Frame_Shift_Del", "Frame_Shift_Ins", "In_Frame_Del", "In_Frame_Ins")) %>% 
  pull(Sample) %>% unique()
length(sample_HLAindel)

## 60 HLA mutations in 52 MSS samples
HLAallMSS <- rbind(maf_tcga, maf_icgc, fill=TRUE) %>% 
  filter(Sample %in% sample_MSS)
sample_HLAMSS <- HLAallMSS  %>% pull(Sample) %>% unique()

## 57 SNVs in 50 samples
sample_HLAsnv <-HLAallMSS %>% filter(!Variant_Classification %in% c("Frame_Shift_Del", "Frame_Shift_Ins", "In_Frame_Del", "In_Frame_Ins")) %>% 
  pull(Sample) %>% unique()
length(sample_HLAsnv)

## 3 Indels in 3 samples
sample_HLAindel <- HLAallMSS %>% filter(Variant_Classification %in% c("Frame_Shift_Del", "Frame_Shift_Ins", "In_Frame_Del", "In_Frame_Ins")) %>% 
  pull(Sample) %>% unique()
length(sample_HLAindel)

## HLA mutations
gritic_path_dict <- ".../Project/MHC_evolution/GRITIC/output/250717/summary250717/dict"
file_list <- list.files(path = gritic_path_dict, pattern = "*mafnongritic.csv", full.names = TRUE)
length(file_list)

gritic_SNV <- list()

for (file in file_list) {
  
  aliquot_id = basename(file)
  aliquot_id = sub("_mafnongritic.csv", "", aliquot_id)
  
  gritic_SNV[[aliquot_id]] <- fread(file) %>% mutate(aliquot_id = aliquot_id)
  
}

# Check -- 9 sample not in 2658 samples and sample_MSS
nonsyn_filter_HLA <- gritic_SNV %>% rbindlist(fill=TRUE) %>% 
  inner_join(., LOH_MSI, by = "aliquot_id") %>% 
  mutate(sample_id = aliquot_id) %>% 
  filter(aliquot_id %in% sample_MSS) 
length(unique(nonsyn_filter_HLA$aliquot_id))

# Check -- 2 sample with only indels
sample_HLAtiming <- nonsyn_filter_HLA %>% filter(Hugo_Symbol %in% c("HLA-A", "HLA-B", "HLA-C"))
length(unique(sample_HLAtiming$aliquot_id))

# Rename coloumns
col_names <- colnames(nonsyn_filter_HLA)
numeric_cols <- grep("^[0-9]+$", col_names)
col_names[numeric_cols] <- paste0("X", col_names[numeric_cols])
colnames(nonsyn_filter_HLA) <- col_names
length(unique(nonsyn_filter_HLA$Tumor_Sample_Barcode))
nonsyn_filter_HLA <- nonsyn_filter_HLA %>% filter(!is.na(X249))
dim(nonsyn_filter_HLA)
length(unique(nonsyn_filter_HLA$aliquot_id))

nonsyn_filter_HLA <- nonsyn_filter_HLA %>% 
  left_join(wgd_median, by = "sample_id") %>%
  mutate(has_NA_in_X = apply(.[, paste0("X", 0:249)], 1, anyNA),
         subclonal_ratio = if_else(has_NA_in_X, NA_real_, apply(.[, paste0("X", 0:249)], 1, function(x) mean(x == 1.01, na.rm = TRUE))),
         lateclonal_ratio = if_else(has_NA_in_X, NA_real_, apply(.[, paste0("X", 0:249)], 1, function(x, wgd) mean(x < 1.01 & x > wgd, na.rm = TRUE), wgd_mediantiming)),
         earlyclonal_ratio = if_else(has_NA_in_X, NA_real_, apply(.[, paste0("X", 0:249)], 1, function(x, wgd) mean(x < wgd, na.rm = TRUE), wgd_mediantiming)),
         timing_gritic = case_when(subclonal_ratio > 0.5 ~ "subclonal",
                                   subclonal_ratio <= 0.5 & !is.na(wgd_mediantiming) & lateclonal_ratio > 0.5 ~ "clonal[late]",
                                   subclonal_ratio <= 0.5 & !is.na(wgd_mediantiming) & earlyclonal_ratio > 0.5 ~ "clonal[early]",
                                   subclonal_ratio <= 0.5 & !is.na(wgd_mediantiming) & lateclonal_ratio <= 0.5 & earlyclonal_ratio <= 0.5 ~ "clonal[NA]",
                                   subclonal_ratio <= 0.5 & is.na(wgd_mediantiming) ~ "clonal[NA]",
                                   has_NA_in_X ~ NA_character_,
                                   TRUE ~ NA_character_))

length(unique(nonsyn_filter_HLA$sample_id))
sample_HLA <- unique(nonsyn_filter_HLA$sample_id)

nonsyn_filter_HLA_check <- nonsyn_filter_HLA %>% filter(sample_id %in% sample_sucess)
length(unique(nonsyn_filter_HLA_check$sample_id))
```

## Update all
```{r}
length(unique(nonsyn_filter_nonPloy$aliquot_id))

nonsyn_filter_nonHLA <- nonsyn_filter_nonPloy %>% 
  filter(!aliquot_id %in% sample_HLA) %>% 
  filter(!Hugo_Symbol %in% c("HLA-A", "HLA-B", "HLA-C"))
length(unique(nonsyn_filter_nonHLA$aliquot_id))

nonsyn_filter <- rbind(nonsyn_filter_nonHLA, nonsyn_filter_HLA, fill=TRUE) %>% 
  filter(aliquot_id %in% sample_MSS)
write.csv(nonsyn_filter, paste0(gritic_path, "/nonsyn_filter.csv"), row.names = F) 

nonsyn_filter <- fread(paste0(gritic_path, "/nonsyn_filter.csv"))
```

# Timeline - Mutations in driver genes
## Generate Background Timing by driver genes
```{r}
timing_driverbackground <- nonsyn_filter %>%
  filter(Hugo_Symbol %in% driver_gene) %>%
  group_by(sample_id) %>%
  summarise(across(all_of(column_names), ~ mean(., na.rm = TRUE)), .groups = "drop") %>%
  mutate(event = "Background_p.A581V", 
         Hugo_Symbol = "Background", 
         Other_Info = NA) %>%
  dplyr::select(c("sample_id", "Hugo_Symbol", "event", everything())) %>% 
  mutate(aliquot_id = sample_id,
         mean_background = rowMeans(as.matrix(dplyr::select(., all_of(column_names)))),
         median_background = rowMedians(as.matrix(dplyr::select(., all_of(column_names))))) %>% 
  filter(!is.na(mean_background))

write_csv(timing_driverbackground, paste0(gritic_path, "/timing_driverbackground.csv"))
```

## Driver genes
### Timing of driver gene 
```{r}
timing_drivergene <- nonsyn_filter %>%
  filter(Hugo_Symbol %in% !!driver_gene) %>%
  group_by(Hugo_Symbol, sample_id) %>%
  summarise(across(column_names, ~ mean(., na.rm = TRUE)), .groups = "drop") %>%
  mutate(event = Hugo_Symbol,
         Other_Info = NA) %>%
  dplyr::select(c("sample_id", "Hugo_Symbol", "event", everything())) %>% 
  mutate(aliquot_id = sample_id,
         mean_timing = rowMeans(as.matrix(dplyr::select(., all_of(column_names)))),
         median_timing = rowMedians(as.matrix(dplyr::select(., all_of(column_names)))),
         pathway = Hugo_Symbol,
         regulator = "Driver") %>% 
  filter(!is.na(mean_timing)) 

write_csv(timing_drivergene, paste0(gritic_path, "/timing_drivergene.csv"))
```

### Deviation of driver genes
```{r}
df_background <- fread(paste0(gritic_path, "/timing_driverbackground.csv")) %>% 
  filter(!is.na(mean_background)) %>%
  mutate(mean_timing = mean_background,
         median_timing = median_background,
         pathway = event,
         regulator = "backgound") %>%
  dplyr::select(c("sample_id", "Hugo_Symbol", "event", "Other_Info", everything()), -mean_background, -median_background)

df_columns <- names(df_background)

df_driver <- fread(paste0(gritic_path, "/timing_drivergene.csv")) %>%
  dplyr::select(c("sample_id", "Hugo_Symbol", "event", "Other_Info", everything()))

df_final <- rbind(df_background, df_driver) 

## Compare the genes with background one by one
diff_driver <- list()

sample_list <- unique(df_final$sample_id)

for (sample in sample_list) {
    
    subset <- df_final %>% filter(sample_id %in% sample) %>% dplyr::select(sample_id, Hugo_Symbol, X0:X249)
    genes <- unique(subset$Hugo_Symbol)
    genes <- genes[genes != "Background"]
    
    if(length(genes) > 0) {
      
      for (gene in genes) {
          
          gene1_data <- subset %>% filter(Hugo_Symbol == !!gene) %>% dplyr::select(X0:X249)
          gene2_data <- subset %>% filter(Hugo_Symbol == "Background") %>% dplyr::select(X0:X249)
          
          diff_values <- gene1_data - gene2_data
          
          pair <- paste0(sample, "_", gene)
          
          print(pair)
          
          diff_driver[[pair]] <- cbind(sample_id = sample, event1 = gene, diff_values)

      }
    }
  }
  
diff_drivergene <- rbindlist(diff_driver, fill = TRUE) 
write.csv(diff_drivergene, file = paste0(gritic_path, "/timing_pair_drivergene.csv"), row.names = F)

diff_drivergene <- fread(paste0(gritic_path, "/timing_pair_drivergene.csv")) %>% 
  mutate(aliquot_id = sample_id,
         pathway = event1,
         regulator = "Driver",
         mean_diff = rowMeans(as.matrix(dplyr::select(., all_of(column_names)))),
         median_diff = rowMedians(as.matrix(dplyr::select(., all_of(column_names))))) %>% 
  left_join(., LOH_type, by = "sample_id")
write_csv(diff_drivergene, paste0(gritic_path, "/diff_drivergene.csv"))

diff_drivergene_simple <- diff_drivergene %>%
  dplyr::select(sample_id, pathway, histology_abbreviation, mean_diff, regulator)
write_csv(diff_drivergene_simple, paste0(gritic_path, "/diff_drivergene_simple.csv"))
```

## All Mutations
### Generate all mutations Timing
```{r}
timing_allmutations <- nonsyn_filter %>% 
  mutate(event = paste0(Hugo_Symbol, "_", POS, "_", HGVSp_Short)) %>%
  group_by(event, sample_id) %>%
  summarise(across(column_names, ~ mean(., na.rm = TRUE)), .groups = "drop") %>%
  mutate(Other_Info = NA,
         Hugo_Symbol = event) %>%
  dplyr::select(c("sample_id", "Hugo_Symbol", "event", everything())) %>% 
  mutate(aliquot_id = sample_id,
         mean_timing = rowMeans(as.matrix(dplyr::select(., all_of(column_names)))),
         median_timing = rowMedians(as.matrix(dplyr::select(., all_of(column_names)))),
         pathway = Hugo_Symbol,
         regulator = "Mutation") %>% 
  filter(!is.na(mean_timing))

write_csv(timing_allmutations, paste0(gritic_path, "/timing_allmutations.csv"))
```

### Deviation of all mutations
```{r}
df_background <- fread(paste0(gritic_path, "/timing_driverbackground.csv")) %>% 
  filter(!is.na(mean_background)) %>%
  mutate(mean_timing = mean_background,
         median_timing = median_background,
         pathway = event,
         regulator = "backgound") %>%
  dplyr::select(c("sample_id", "Hugo_Symbol", "event", "Other_Info", everything()), -mean_background, -median_background)

df_columns <- names(df_background)

df_driver <- fread(paste0(gritic_path, "/timing_allmutations.csv")) %>% mutate(Hugo_Symbol = event) %>%
  dplyr::select(c("sample_id", "Hugo_Symbol", "event", "Other_Info", everything())) %>%
  dplyr::select(all_of(df_columns))

sample_background <- unique(df_background$sample_id)
sample_mut <- unique(df_driver$sample_id)
sample_comm <- intersect(sample_background, sample_mut)

df_final <- rbind(df_background, df_driver) %>% filter(sample_id %in% sample_comm) %>%
  distinct()

## Compare the genes with background one by one
diff_driver <- list()

sample_list <- unique(df_final$sample_id)

for (sample in sample_list) {
    
    subset <- df_final %>% filter(sample_id %in% sample) %>% dplyr::select(sample_id, Hugo_Symbol, X0:X249)
    genes <- unique(subset$Hugo_Symbol)
    genes <- genes[genes != "Background"]
    
    if(length(genes) > 0) {
      
      for (gene in genes) {
          
          gene1_data <- subset %>% filter(Hugo_Symbol == !!gene) %>% dplyr::select(X0:X249)
          gene2_data <- subset %>% filter(Hugo_Symbol == "Background") %>% dplyr::select(X0:X249)
          
          diff_values <- gene1_data - gene2_data
          
          pair <- paste0(sample, "_", gene)
          
          print(pair)
          
          diff_driver[[pair]] <- cbind(sample_id = sample, event1 = gene, diff_values)

      }
    }
  }
  
diff_allmutations <- rbindlist(diff_driver, fill = TRUE) 
write.csv(diff_allmutations, file = paste0(gritic_path, "/timing_pair_allmutations.csv"), row.names = F)

diff_allmutations <- fread(paste0(gritic_path, "/timing_pair_allmutations.csv")) %>%
  mutate(aliquot_id = sample_id,
         pathway = event1,
         regulator = "Driver",
         mean_diff = rowMeans(as.matrix(dplyr::select(., all_of(column_names)))),
         median_diff = rowMedians(as.matrix(dplyr::select(., all_of(column_names))))) %>% 
    left_join(., LOH_MSI, by = "aliquot_id")
write_csv(diff_allmutations, paste0(gritic_path, "/diff_allmutations.csv"))

diff_allmutations_simple <- diff_allmutations %>% 
  dplyr::select(sample_id, pathway, histology_abbreviation, mean_diff, regulator)
write_csv(diff_drivergene_simple, paste0(gritic_path, "/diff_allmutations_simple.csv"))
```

### all genes
```{r}
cutoff_mut = 0.6
cutoff_pathway = 0.5
cutoff_ratio = 2

data <- fread(paste0(gritic_path, "/diff_allmutations.csv")) %>% 
  mutate(Hugo_Symbol = sub("_.*", "", event1),
         sample_id = aliquot_id) %>%
  group_by(sample_id, Hugo_Symbol) %>%
  summarise(across(all_of(column_names), ~ mean(., na.rm = TRUE)), .groups = "drop") %>%
  mutate(pathway = Hugo_Symbol,
         early_ratio = rowMeans(dplyr::select(., all_of(column_names)) < 0, na.rm = TRUE),
         late_ratio = rowMeans(dplyr::select(., all_of(column_names)) > 0, na.rm = TRUE),
         undetemined_ratio = rowMeans(dplyr::select(., all_of(column_names)) == 0, na.rm = TRUE),
         timing_cat = case_when(early_ratio > cutoff_mut ~ "Early", 
                                late_ratio > cutoff_mut ~ "Late", 
                                TRUE ~ "Undetermined"),
         aliquot_id = sample_id) %>%
  inner_join(type_list, by = "aliquot_id") %>%
  dplyr::select(c("sample_id", "aliquot_id", "histology_abbreviation", "Hugo_Symbol", "pathway", everything()))

write.csv(data,  paste0(gritic_path, "/diff_allgene.csv"), row.names = F)
```

## Driver mutations
### Deviation of driver mutations
```{r}
df_background <- fread(paste0(gritic_path, "/timing_driverbackground.csv")) %>% 
  filter(!is.na(mean_background)) %>%
  mutate(mean_timing = mean_background,
         median_timing = median_background,
         pathway = event,
         regulator = "backgound") %>%
  dplyr::select(c("sample_id", "Hugo_Symbol", "event", "Other_Info", everything()), -mean_background, -median_background)

df_drivermutation <- fread(paste0(gritic_path, "/timing_drivermutation240806.csv")) %>%
  dplyr::select(c("sample_id", "Hugo_Symbol", "event", "Other_Info", everything()))

df_final <- rbind(df_background, df_drivermutation)
# write.csv(df_final, file = paste0(gritic_path, "/", "pair_APC.csv"), row.names = F)

## Compare the genes with background one by one
diff_driver <- list()

sample_list <- unique(df_final$sample_id)

for (sample in sample_list) {
    
    subset <- df_final %>% filter(sample_id %in% sample) %>% dplyr::select(sample_id, Hugo_Symbol, X0:X249)
    genes <- unique(subset$Hugo_Symbol)
    genes <- genes[genes != "Background"]
    
    if(length(subset$Hugo_Symbol) > 1) {
      
      for (gene in genes) {
          
          gene1_data <- subset %>% filter(Hugo_Symbol == !!gene) %>% dplyr::select(X0:X249)
          gene2_data <- subset %>% filter(Hugo_Symbol == "Background") %>% dplyr::select(X0:X249)
          
          diff_values <- gene1_data - gene2_data
          
          pair <- paste0(sample, "_", gene)
          
          print(pair)
          
          diff_driver[[pair]] <- cbind(sample_id = sample, event1 = gene, diff_values)

      }
    }
  }
  
diff_drivermutation <- rbindlist(diff_driver, fill = TRUE) 
write.csv(diff_drivermutation, file = paste0(gritic_path, "/timing_pair_drivermutation.csv"), row.names = F)

diff_drivermutation <- fread(paste0(gritic_path, "/timing_pair_drivermutation.csv")) %>% 
  mutate(aliquot_id = sample_id,
         pathway = event1,
         regulator = "Driver",
         mean_diff = rowMeans(as.matrix(dplyr::select(., all_of(column_names)))),
         median_diff = rowMedians(as.matrix(dplyr::select(., all_of(column_names))))) %>% 
  left_join(., LOH_type, by = "sample_id") %>% 
  #filter(sample_id %in% sample_MSS) %>%
  dplyr::select(sample_id, pathway, histology_abbreviation, mean_diff, regulator)
write_csv(diff_drivermutation, paste0(gritic_path, "/diff_drivermutation.csv"))
```

## APM timing
```{r}
APM_timing <- nonsyn_filter %>% filter(Hugo_Symbol %in% !!APM) %>% 
  filter(!is.na(X0)) %>% 
  filter(sample_id %in% sample_MSS) %>% 
  filter(sample_id %in% sample_sucess)
write.csv(APM_timing, paste0(gritic_path, "/Timing_APM.csv"), row.names = F)

dim(APM_timing)
table(APM_timing$timing_gritic)

APM_sample <- unique(APM_timing$sample_id)
length(APM_sample)

df_APM <- APM_timing %>% filter(aliquot_id %in% APM_sample) %>%
  filter(Hugo_Symbol %in% !!APM) %>%
  group_by(sample_id) %>%
  summarise(across(column_names, ~ mean(., na.rm = TRUE)), .groups = "drop")  %>%
  mutate(event = "APM_p.A581V", 
         Hugo_Symbol = "APM", 
         Other_Info = NA,
         aliquot_id = sample_id) %>%
  dplyr::select(c("sample_id", "Hugo_Symbol", "event", "Other_Info", "aliquot_id", everything()))

df_driver <- fread(paste0(gritic_path, "/timing_driverbackground.csv")) %>%
  filter(aliquot_id %in% APM_sample) %>%
  mutate(event = "Mean_p.A581V", 
         Hugo_Symbol = "Mean", 
         Other_Info = NA) %>%
  dplyr::select(c("sample_id", "Hugo_Symbol", "event", "Other_Info", "aliquot_id", everything()), -mean_background, -median_background)

final_df <- rbind(df_APM, df_driver)
write.csv(final_df, paste0(gritic_path, "/APM_pair.csv"), row.names = F)

diff_APM <- list()

for (sample in APM_sample) {
  
  subset_APM_one <- final_df %>% filter(sample_id %in% sample) %>% dplyr::select(sample_id, Hugo_Symbol, X0:X249)
  genes <- unique(subset_APM_one$Hugo_Symbol)
  
  if(length(genes) > 1) {
    
  for (i in 1:(length(genes) - 1)) {
    for (j in (i + 1):length(genes)) {
      
      gene1_data <- subset_APM_one %>% filter(Hugo_Symbol == genes[i]) %>% dplyr::select(X0:X249)
      gene2_data <- subset_APM_one %>% filter(Hugo_Symbol == genes[j]) %>% dplyr::select(X0:X249)
      
      diff_values <- gene1_data - gene2_data
      
      pair <- paste0(sample, "_", genes[i], "_", genes[j])
      
      print(pair)
      
      diff_APM[[pair]] <- cbind(sample_id = sample, event1 = genes[i], event2 = genes[j], diff_values)
      
    }
    }
    }
}

APM_driver <- rbindlist(diff_APM, fill = TRUE) %>% 
  mutate(aliquot_id = sample_id,
         mean_diff = rowMeans(as.matrix(dplyr::select(., all_of(column_names)))),
         median_diff = rowMedians(as.matrix(dplyr::select(., all_of(column_names))))) %>% 
  inner_join(., LOH_type, by = "sample_id") 
write.csv(APM_driver, file = paste0(gritic_path, "/", "/timing_pair_APM.csv"), row.names = F)

diff_APM <- fread(paste0(gritic_path, "//timing_pair_APM.csv")) %>% 
  mutate(pathway = "APM",
         regulator = "APM")
write.csv(diff_APM, file = paste0(gritic_path, "/diff_APM.csv"), row.names = F)

diff_APM_simple <- diff_APM %>%
  dplyr::select(sample_id, pathway, histology_abbreviation, mean_diff, regulator)
write_csv(diff_APM_simple, paste0(gritic_path, "/diff_APM_simple.csv"))
```

## Deviation of WGD timing
```{r}
timing_wgd <- fread(".../Project/MHC_evolution/GRITIC/output/wgd_median.csv")
timing_driverbackground <- fread(paste0(gritic_path, "/timing_driverbackground.csv"))

diff_wgd <- timing_driverbackground %>% 
  inner_join(timing_wgd, by = "sample_id") %>% 
  mutate(mean_diff = wgd_mediantiming -  mean_background,
         pathway = "WGD",
         regulator = "WGD") %>% 
  filter(sample_id %in% sample_MSS) %>%
  dplyr::select(-matches("\\.y$"), -Sample) %>%
  rename_with(~ str_replace_all(., "\\.x$", "")) %>%
  dplyr::select(sample_id, pathway, histology_abbreviation, mean_diff, regulator)

write_csv(diff_wgd, paste0(gritic_path, "/diff_wgd.csv"))
```

## Deviation of pathway timing
### Timing difference
```{r}
df_driver <- fread(paste0(gritic_path, "/timing_driverbackground.csv")) %>%
             dplyr::select(c("sample_id", "Hugo_Symbol", "event", "Other_Info", everything())) %>% 
             dplyr::select(-aliquot_id, -mean_background, -median_background)

cell_types <- c("Tcells", "MHCregulators", "NKcells", "Macrophages", "GammadeltaTcells")
regulators <- c("pos", "neg")

for (cell_type in cell_types) {
  
   CRISPR_path <- paste0(".../Project/CRISPR/CRISPRscreen20240118/", cell_type)
   
   for (reg in regulators) {
    
     CRISPR_top10 <- read.csv(paste0(CRISPR_path, "/Summary/", cell_type, "_C5BP_", reg, "_list.csv"), header = T) %>% 
       pull(Description)
     
     CRISPR_pathways <- fread(paste0(CRISPR_path, "/Summary/", cell_type, "_C5BP_", reg, "_allgenelist.csv")) %>% 
       filter(Pathway1 %in% CRISPR_top10) %>%
       dplyr::select(Gene, Pathway1)
     
     pathway_list <- unique(CRISPR_pathways$Pathway1)
     
     #save_path <- paste0(CRISPR_path, "/Summary/TimingSNV/timing_ttest/drivermutation/250514/")
     
     dir.create(save_path, recursive = TRUE)
        
     for (pathway in pathway_list) {
       
        pathway_title <- str_to_title(pathway)
        plot_key <- paste0(cell_type, "_", reg, "_", pathway)
        
        CRISPR_gene <- CRISPR_pathways %>% filter(Pathway1 == !!pathway) %>% pull(Gene)
          
          tryCatch({
          
           Reg_timing <- nonsyn_filter %>% filter(Hugo_Symbol %in% CRISPR_gene) %>% filter(!is.na(X0))
           Reg_sample <- unique(Reg_timing$sample_id)

           df_Reg <- Reg_timing %>% filter(sample_id %in% Reg_sample) %>%
             filter(Hugo_Symbol %in% !!CRISPR_gene) %>%
             group_by(sample_id) %>%
             summarise(across(column_names, ~ mean(., na.rm = TRUE)), .groups = "drop") %>%
             mutate(event = "Reg_nonsyn", 
                    Hugo_Symbol = "Reg", 
                    Other_Info = NA) %>%
             dplyr::select(c("sample_id", "Hugo_Symbol", "event", "Other_Info", everything()))
 
           df_final <- rbind(df_Reg, df_driver)
           
           df_pair <- list()
           diff_driver <- list()
           reg_driver <- list()
           
           sample_list <- unique(df_final$sample_id)
         
           for (sample in sample_list) {
    
           subset <- df_final %>% filter(sample_id %in% !!sample) %>% dplyr::select(sample_id, Hugo_Symbol, X0:X249)
           genes <- unique(subset$Hugo_Symbol)
           # genes <- genes[genes != "Background"]
    
           if(length(genes) > 1) {
            
             for (gene in genes) {
               
               if(gene != "Background") {
          
               gene1_data <- subset %>% filter(Hugo_Symbol == !!gene) %>% dplyr::select(X0:X249)
               gene2_data <- subset %>% filter(Hugo_Symbol == "Background") %>% dplyr::select(X0:X249)
          
               diff_values <- gene1_data - gene2_data
               pair <- paste0(sample, "_", gene)
               print(pair)
          
               diff_driver[[pair]] <- cbind(sample_id = sample, event1 = gene, diff_values)
             }
             }
             }
           }
           
           final_driver <- rbindlist(diff_driver, fill = TRUE) %>% 
            dplyr::select(sample_id, column_names) %>% 
            mutate(Hugo_Symbol = "Difference",
                   event = "Difference",
                   Other_Info = NA) %>% 
            right_join(LOH_type,., by = "sample_id") %>%
            mutate(mean_diff = rowMeans(as.matrix(dplyr::select(., all_of(column_names)))),
                   median_diff = rowMedians(as.matrix(dplyr::select(., all_of(column_names))))) %>%
            arrange(histology_abbreviation, sample_id)
           write.csv(final_driver, file = paste0(save_path, "/difference_", cell_type, "_", reg, "_", pathway, ".csv"), row.names = F)
           
           rm(final_driver)
           rm(reg_driver)
           
             }, error = function(e) {
                    message("Error processing ", cell_type, "_", reg, "_", pathway, ": ", e$message)
             })
     }
   }
}
```

### Summarize results
```{r}
rename_list <- fread(".../Project/CRISPR/CRISPRscreen20240118/rename_list.csv") %>% 
  mutate(pathway = str_to_title(pathway),
         pathway = sub("/", "_", pathway))

cell_types <- c("Tcells", "MHCregulators", "NKcells", "Macrophages", "GammadeltaTcells")
data_list <- list()

for (cell_type in cell_types) {
   
    CRISPR_path <- paste0(".../Project/CRISPR/CRISPRscreen20240118/", cell_type)
    timing_path <- paste0(CRISPR_path, "/Summary/TimingSNV/timing_ttest/drivermutation/250514/")
   
    file_paths <- list.files(timing_path, pattern = "^difference_", full.names = TRUE, recursive = TRUE)
    file_paths <- grep("\\.csv$", file_paths, value = TRUE)

    for (file in file_paths) {
       
     file_name <- basename(file)
     file_name <- sub("difference_", "", file_name)
     file_name <- sub(".csv", "", file_name)
     data_list[[file_name]] <- fread(file) %>% mutate(Study = file_name)
    }
 }

diff_pathway <- tidy_list(data_list) %>% 
  tidyr::separate(Study, into = c("celltype", "regulator", "pathway"), extra = "merge", fill = "right")  %>%
  mutate(pathway = str_to_title(pathway)) %>% 
  inner_join(rename_list,., by = "pathway") %>% 
  mutate(pathway = pathway_new)
write_csv(diff_pathway, paste0(gritic_path, "/diff_pathway_all.csv"))

diff_pathway <- diff_pathway %>% dplyr::select(sample_id, pathway, histology_abbreviation, mean_diff, regulator, celltype)
write_csv(diff_pathway, paste0(gritic_path, "/diff_pathway.csv"))
```

## Combine all
```{r}
cutoff_mut = 0.6
cutoff_pathway = 0.5
cutoff_ratio = 2

## Pathway
diff_pathway <- fread(paste0(gritic_path, "/diff_pathway_all.csv")) %>% 
  filter(sample_id %in% sample_MSS) %>%
  mutate(aliquot_id = sample_id) %>%
  dplyr::select(all_of(column_names), pathway, histology_abbreviation, aliquot_id, mean_diff, median_diff,
  histology_abbreviation, celltype, regulator)

## APM
diff_APM <- fread(paste0(gritic_path, "//timing_pair_APM.csv")) %>% 
  filter(sample_id %in% sample_MSS) %>%
  mutate(pathway = "APM", 
         regulator = "APM",
         celltype = "other") %>% 
  dplyr::select(all_of(column_names), "pathway", "histology_abbreviation", "aliquot_id", "mean_diff", "median_diff", "celltype", "regulator") 

## Driver Genes
diff_drivergene <- fread(paste0(gritic_path, "/timing_pair_drivergene.csv")) %>% 
  filter(sample_id %in% sample_MSS) %>%
  left_join(., LOH_type, by = "sample_id") %>% 
  mutate(aliquot_id = sample_id,
         regulator = "Driver",
         pathway = event1,
         celltype = "other",
         mean_diff = rowMeans(as.matrix(dplyr::select(., all_of(column_names)))),
         median_diff = rowMedians(as.matrix(dplyr::select(., all_of(column_names))))) %>% 
  dplyr::select(all_of(column_names), "pathway", "histology_abbreviation", "aliquot_id", "mean_diff", "median_diff", "celltype", "regulator") 

## Combine with driver genes
diff_all <- rbind(diff_pathway, diff_APM) %>% rbind(., diff_drivergene) %>% filter(aliquot_id %in% sample_MSS) %>% 
       mutate(early_ratio = rowMeans(dplyr::select(., all_of(column_names)) < 0, na.rm = TRUE),
              late_ratio = rowMeans(dplyr::select(., all_of(column_names)) > 0, na.rm = TRUE),
              undetemined_ratio = rowMeans(dplyr::select(., all_of(column_names)) == 0, na.rm = TRUE),
              timing_cat = case_when(early_ratio > cutoff_mut ~ "Early", 
                                     late_ratio > cutoff_mut ~ "Late", 
                                     TRUE ~ "Undetermined"),
               p20 = apply(dplyr::select(., all_of(column_names)), 1, function(row) quantile(row, 0.2, na.rm = TRUE)),
               p80 = apply(dplyr::select(., all_of(column_names)), 1, function(row) quantile(row, 0.8, na.rm = TRUE)),
               regulator = case_when(regulator == "neg" ~ "Neg",
                                     regulator == "pos" ~ "Pos",
                                     regulator == "both" ~ "Both",
                                     TRUE ~ regulator),
              pathway = case_when(pathway == "Antigen Presentation via MHC I" & celltype == "MHCregulators" ~ paste0(pathway, " (MHC-I) "),
                                  pathway == "Antigen Presentation via MHC I" & celltype == "Tcells" ~ paste0(pathway, " (Tcells) "),
                                  pathway == "Antigen Presentation via MHC Ib" & celltype == "NKcells" ~ paste0(pathway, " (NKcells) "),
                                  pathway == "IFNg Pathway" & celltype == "NKcells" ~ paste0(pathway, " (NKcells) "),
                                  pathway == "IFNg Signaling" & celltype == "MHCregulators" ~ paste0(pathway, " (MHC-I)"),
                                  pathway == "IFNg Signaling" & celltype == "Tcells" ~ paste0(pathway, " (Tcells)"),
                                  pathway == "Regulation of Autophagy" & celltype == "MHCregulators" ~ paste0(pathway, " (MHC-I)"),
                                  pathway == "Autophagy" & celltype == "Tcells" ~ paste0(pathway, " (Tcells)"),
                                  pathway == "HSWI SNF Pathway" ~ "SWI/SNF Pathway",
                                  pathway == "Corticosteroid Receptor Signaling Pathway" ~ "Corticosteroid Receptor Signaling",
                                  pathway == "PD-L1-PD-1 Checkpoint Pathway In Cancer" ~ "PD-L1/PD-1 Checkpoint Pathway",
                                  pathway == "NOD1_2 Signaling Pathway" ~ "NOD1/2 Signaling Pathway",
              TRUE ~ pathway)) %>% filter(pathway != "WGD" & pathway != "Chromatin Disassembly")
write_csv(diff_all, paste0(gritic_path, "/diff_alldrivergenes.csv"))
```

# All Pathway -- Propotion for all
```{r}
cutoff_mut = 0.6
cutoff_pathway = 0.5
cutoff_ratio = 2
#path_timeline = ".../Project/CRISPR/CRISPRscreen20240118/timeline/241015/"
path_timeline = ".../Project/MHC_evolution/GRITIC/output/250717/summary250717/timeline/"

dir.create(path_timeline)

diff_all <- fread(paste0(gritic_path, "/diff_alldrivergenes.csv")) %>% 
       mutate(early_ratio = rowMeans(dplyr::select(., all_of(column_names)) < 0, na.rm = TRUE),
              late_ratio = rowMeans(dplyr::select(., all_of(column_names)) > 0, na.rm = TRUE),
              undetemined_ratio = rowMeans(dplyr::select(., all_of(column_names)) == 0, na.rm = TRUE),
              timing_cat = case_when(early_ratio > cutoff_mut ~ "Early", 
                                     late_ratio > cutoff_mut ~ "Late", 
                                     TRUE ~ "Undetermined"),
               p20 = apply(dplyr::select(., all_of(column_names)), 1, function(row) quantile(row, 0.2, na.rm = TRUE)),
               p80 = apply(dplyr::select(., all_of(column_names)), 1, function(row) quantile(row, 0.8, na.rm = TRUE)),
               regulator = case_when(regulator == "neg" ~ "Neg",
                                     regulator == "pos" ~ "Pos",
                                     regulator == "both" ~ "Both",
                                     TRUE ~ regulator),
              pathway = case_when(pathway == "Antigen Presentation via MHC-I" & celltype == "MHCregulators" ~ paste0(pathway, " (MHC-I) "),
                                  pathway == "Antigen Presentation via MHC-I" & celltype == "Tcells" ~ paste0(pathway, " (Tcells) "),
                                  pathway == "Antigen Presentation via MHC-Ib" & celltype == "NKcells" ~ paste0(pathway, " (NKcells) "),
                                  pathway == "IFNg Pathway" & celltype == "NKcells" ~ paste0(pathway, " (NKcells) "),
                                  pathway == "IFNg Signaling" & celltype == "MHCregulators" ~ paste0(pathway, " (MHC-I)"),
                                  pathway == "IFNg Signaling" & celltype == "Tcells" ~ paste0(pathway, " (Tcells)"),
                                  pathway == "Necroptosis" & celltype == "Tcells" ~ paste0(pathway, " (Tcells)"),
                                  pathway == "Necroptosis" & celltype == "NKcells" ~ paste0(pathway, " (NKcells)"),
                                  pathway == "Regulation of Autophagy" & celltype == "MHCregulators" ~ paste0(pathway, " (MHC-I)"),
                                  pathway == "Autophagy" & celltype == "Tcells" ~ paste0(pathway, " (Tcells)"),
                                  pathway == "HSWI SNF Pathway" ~ "SWI/SNF Pathway",
                                  pathway == "Corticosteroid Receptor Signaling Pathway" ~ "Corticosteroid Receptor Signaling",
                                  pathway == "PD-L1 PD-1 Checkpoint Pathway in Cancer" ~ "PD-L1/PD-1 Checkpoint Pathway",
                                  pathway == "NOD1_2 Signaling Pathway" ~ "NOD1/2 Signaling Pathway",
                                  pathway == "TNF Pathway" & celltype == "Tcells" ~ paste0(pathway, " (Tcells)"),
                                  pathway == "TNF Pathway" & celltype == "NKcells" ~ paste0(pathway, " (NKcells)"),
              TRUE ~ pathway)) %>% filter(pathway != "WGD" & pathway != "Chromatin Disassembly", pathway != "KIT Pathway") %>% filter(aliquot_id %in% sample_MSS)

write_csv(diff_all, paste0(gritic_path, "/diff_all.csv"))
saveRDS(diff_all, file = "/Users/wchen20/Desktop/Website/Shiny/EvoGIE/data/Cohorts/PCAWG/diff_all_PCAWG.rds")
```

## Plot Bar
```{r}
cutoff_mut = 0.6
cutoff_pathway = 0.5
cutoff_ratio = 2
mode = 2
num_filter = 4
  
diff_all <- fread(paste0(gritic_path, "/diff_all.csv")) %>% filter(aliquot_id %in% sample_MSS, aliquot_id %in% sample_sucess)

pathway_list <- diff_all %>% distinct(pathway, regulator, .keep_all = F)

types <- unique(diff_all$histology_abbreviation)

for (type in types) {
  
  title = study = type
  
  # Combine and select
  drivergene_cancer <- driver_list %>% filter(grepl(type, Tissue)) %>% pull(Gene)
  data_driver <- diff_all %>% filter(histology_abbreviation == !!type, pathway %in% drivergene_cancer) %>% filter(!is.na(mean_diff)) 
  data_nondriver <- diff_all %>% filter(histology_abbreviation == !!type, regulator != "Driver") %>% filter(!is.na(mean_diff)) 
  
  data_all <- rbind(data_driver, data_nondriver)
  
  data_filter1 <- data_all %>% group_by(pathway, histology_abbreviation) %>%
    summarise(count = n()) %>%
    ungroup() %>%
    inner_join(., data_all, by = c("histology_abbreviation", "pathway")) %>%
    mutate(n_base = 1,
           percent_base = n_base/count,
           n_late = 250 * late_ratio,
           n_undetermined = 250 * undetemined_ratio,
           ratio_plot = (n_late + n_undetermined/2) / 250,
           timing_bin = cut(ratio_plot, breaks = range_breaks, include.lowest = TRUE)) %>% 
    arrange(desc(count))
  
  data_filter <- data_filter1 %>% filter(timing_cat != "Undetermined") %>%
    group_by(pathway) %>% 
    summarise(Sum_determined = n()) %>% 
    ungroup %>% 
    inner_join(data_filter1, ., by = "pathway") %>%
    filter(Sum_determined > 4)
  
  height = 7
  
  if (type == "Liver-HCC") {
    
    data_filter <- data_filter %>% filter(!pathway %in% c("AXIN1", "CTNNB1", "SETDB1", "APOB", "NFE2L2", "ARID2", "MACF1", "ASH1L"))
  
  }  else if (type %in% c("Breast-AdenoCA", "Panc-AdenoCA")) {
      
    data_filter <- data_filter %>% filter(!pathway %in% "Autophagy (Tcells)")
    
  } else {
    
    data_filter <- data_filter
    
  }
  
   write_csv(data_filter, paste0(path_timeline, "/filter_", type, ".csv"))
  #data_filter <- fread(paste0(path_timeline, "/filter_", type, ".csv"))
  
  Freq <- as.data.frame(table(data_filter$pathway, data_filter$timing_cat))
  
  if (nrow(Freq) > 2 ) {
    
    Freq <- Freq %>% setNames(c("pathway", "time_period", "Freq")) 
    type_sum <- as.data.frame(table(LOH_MSI$histology_abbreviation)) %>% filter(Var1 == !!type) %>% pull(Freq)
    
    Sum <- as.data.frame(table(data_filter$pathway)) %>% 
      mutate(Type_sum = type_sum) %>% 
      setNames(c("pathway", "Sum", "Sum_type")) %>% 
      mutate(Prevelance = Sum/Sum_type) 
    
    data_prop <- inner_join(Freq, Sum, by = "pathway") %>% mutate(Proportion = Freq/Sum) %>% inner_join(pathway_list, by = "pathway")
    
    if (nrow(data_filter) > 1) {
      
      ratio <- data_prop %>% filter(time_period != "Undetermined") %>%
        group_by(pathway, time_period) %>% 
        summarise(sum_freq = sum(Freq), .groups = 'drop') %>% 
        pivot_wider(names_from = time_period, values_from = sum_freq) %>% 
        mutate(ratio = (Late + 1)/(Early + 1)) %>% 
        inner_join(., data_prop, by = "pathway") %>%
        distinct(pathway, .keep_all = TRUE) %>% 
        mutate(pro_early = Early/Sum,
               pro_late = Late/Sum) %>% 
        arrange(desc(ratio)) 
      
       write_csv(ratio, paste0(path_timeline, "/ratio_", type, ".csv"))
       
       order_all <- ratio %>% arrange(desc(ratio)) %>% pull(pathway)
      
       prop_early <- data_prop %>% filter(time_period == "Early", Proportion >= cutoff_pathway) %>% arrange(desc(Proportion), pathway) %>% pull(pathway) %>% as.character() %>% unique()
       prop_late <- data_prop %>% filter(time_period == "Late", Proportion >= cutoff_pathway) %>% arrange(desc(Proportion), pathway) %>% pull(pathway) %>% as.character() %>% unique()
       prop_undetermined <- data_prop %>% filter(!pathway %in% c(prop_early, prop_late), time_period == "Early") %>% arrange(desc(Proportion), pathway) %>% pull(pathway) %>% as.character() %>% unique()
      
       ratio_early <- ratio %>% filter(ratio < 1/cutoff_ratio) %>% arrange(ratio) %>% pull(pathway) %>% as.character() %>% unique()
       ratio_late <- ratio %>% filter(ratio > cutoff_ratio) %>% arrange(ratio) %>% pull(pathway) %>% as.character() %>% unique()
       ratio_undetermined <- ratio %>% filter(!pathway %in% c(ratio_early, ratio_late)) %>% arrange(ratio, desc(pathway)) %>% pull(pathway) %>% as.character() %>% unique()
      
      ratio_earlydriver <- ratio %>% filter(ratio < 1/cutoff_ratio, regulator == "Driver") %>% arrange(ratio) %>% pull(pathway) %>% as.character() %>% unique()
      ratio_latedriver <- ratio %>% filter(ratio > cutoff_ratio, regulator == "Driver") %>% arrange(ratio) %>% pull(pathway) %>% as.character() %>% unique()
      ratio_undetermineddriver <- ratio %>% filter(!pathway %in% c(ratio_early, ratio_late), regulator == "Driver") %>% arrange(ratio, desc(pathway)) %>% pull(pathway) %>% as.character() %>% unique()
      
      ratio_earlypathway <- ratio %>% filter(ratio < 1/cutoff_ratio, regulator != "Driver") %>% arrange(ratio) %>% pull(pathway) %>% as.character() %>% unique()
      ratio_latepathway <- ratio %>% filter(ratio > cutoff_ratio, regulator != "Driver") %>% arrange(ratio) %>% pull(pathway) %>% as.character() %>% unique()
      ratio_undeterminedpathway <- ratio %>% filter(!pathway %in% c(ratio_early, ratio_late), regulator != "Driver") %>% arrange(ratio, desc(pathway)) %>% pull(pathway) %>% as.character() %>% unique()
      
      if (mode == 1) {
        
        order_early <- prop_early
        order_late <- prop_late
        order_undetermined <- prop_undetermined
        
      }
      
      if (mode == 2) {
        
        order_early <- ratio_early
        order_late <- ratio_late
        order_undetermined <- ratio_undetermined
        
        order_earlydriver <- ratio_earlydriver
        order_latedriver <- ratio_latedriver
        order_undetermineddriver <- ratio_undetermineddriver
        
        order_earlypathway <- ratio_earlypathway
        order_latepathway <- ratio_latepathway
        order_undeterminedpathway <- ratio_undeterminedpathway
        
        order_all <- ratio %>% arrange(desc(ratio)) %>% pull(pathway)
      }
      
      if (mode == 3) {
        
        order_early <- intersect(prop_early, ratio_early)
        order_late <- intersect(prop_late, ratio_late)
        order_undetermined <- setdiff(order_all, union(order_early, order_late))
        
      }
      
      ## Plot for proportion
      data_bar <- data_filter %>% filter(!is.na(late_ratio))  %>% 
        mutate(pathway = factor(pathway, levels = order_all),
               timing_bin = factor(timing_bin, levels = rev(timing_bin_levels)))
      
      plot_bar <- ggplot(data_bar, aes(x = pathway, fill = timing_bin)) +
        geom_bar(position = "fill") +
         theme_minimal() +
        labs(x = "",  y = "Sample", fill = "Timing Probability", title = "") +
        scale_fill_manual(values = color_vector, drop = FALSE) +
        theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.5, size = 18, colour = "black"),
              axis.text.y = element_text(size = 16, colour = "black", hjust = 1),
              plot.title = element_text(size = 20, vjust = 0.5),
              axis.title.y = element_text(size = 18),
              axis.title.x = element_text(size = 18),
              legend.text = element_blank(),  
              legend.title = element_text(size = 16),
              legend.position = "none",
              plot.margin = margin(0, 30, 5, 5), 
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank()) +
        coord_flip()
      
      plot_bar_legend <- cowplot::plot_grid(plot_bar, legend, ncol = 2, rel_widths = c(1, 0.1))
      ggsave(filename = paste0(path_timeline, "/Barplot_", type, ".pdf"), plot = plot_bar_legend, width = 9, height = 7.5, units = "in")
    
      ratio_row <- ratio %>%
        mutate(timing_cat = case_when(
          ratio > 2 ~ "Late",
          ratio < 0.5 ~ "Early",
          TRUE ~ "Undetermined"),
          `Ratio(Log2)` = log2(ratio),
          pathway = factor(pathway, levels = order_all))
      
      max <- max(abs(ratio_row$`Ratio(Log2)`))
      
      plot_ratio <- ggplot(ratio_row, aes(x = `Ratio(Log2)`, y = pathway, fill = timing_cat)) +
        geom_point(size = 3, shape = 21, color = "black") +  # shape 21 allows for fill and color
        theme_bw() +
        theme(axis.text.y = element_blank(),
              axis.title.y = element_blank(),
              axis.title.x = element_text(size = 18),
              legend.title = element_blank(), 
              legend.text = element_text(size = 12),
              axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.5, size = 18, colour = "black"),
              plot.margin = margin(18, 30, 5, 5),
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank()) +
        scale_fill_manual(values = color_bin) +  
        xlim(-(max + 0.5), max + 0.5) + 
        geom_vline(xintercept = 1, linetype = "dashed", color = "grey60") +  
        geom_vline(xintercept = -1, linetype = "dashed", color = "grey60")

      plot_barcom <- cowplot::plot_grid(plot_bar, 
      legend, 
      plot_ratio,
      ncol = 3,  
      rel_widths = c(1, 0.1, 0.5))

      order_earlydriver <- ratio_earlydriver
      order_latedriver <- ratio_latedriver
      order_undetermineddriver <- ratio_undetermineddriver
      
      early_textdriver <- paste("Driver gene\n", paste(order_earlydriver, collapse = "\n"))
      late_textdriver <- paste("Driver gene\n", paste(order_latedriver, collapse = "\n"))
      variable_textdriver <- paste("Driver gene\n", paste(order_undetermineddriver, collapse = "\n"))

      early_textpathway <- paste("Pathway\n", paste(order_earlypathway, collapse = "\n"))
      late_textpathway <- paste("Pathway\n", paste(order_latepathway, collapse = "\n"))
      variable_textpathway <- paste("Pathway\n", paste(order_undeterminedpathway, collapse = "\n"))

      if (length(order_earlydriver) > 0  & length(order_earlypathway) > 0) {
        early_text <- paste(early_textdriver, "\n", early_textpathway)
      } else if (length(order_earlydriver) == 0 & length(order_earlypathway) > 0) { 
        early_text <- early_textpathway
      } else if (length(order_earlydriver) > 0 & length(order_earlypathway) == 0) { 
        early_text <- early_textdriver
      } else {
        early_text <- "" 
      }
      
      if (length(order_latedriver) > 0  & length(order_latepathway) > 0) {
        late_text <- paste(late_textdriver, "\n", late_textpathway)
      } else if (length(order_latedriver) == 0 & length(order_latepathway) > 0) { 
        late_text <- late_textpathway
      } else if (length(order_latedriver) > 0 & length(order_latepathway) == 0) { 
        late_text <- late_textdriver
      } else {
        late_text <- "" 
           }
      
      if (length(order_undetermineddriver) > 0  & length(order_undeterminedpathway) > 0) {
        variable_text <- paste(variable_textdriver, "\n", variable_textpathway)
      } else if (length(order_undetermineddriver) == 0 & length(order_undeterminedpathway) > 0) { 
        variable_text <- variable_textpathway
      } else if (length(order_undetermineddriver) > 0 & length(order_undeterminedpathway) == 0) { 
        variable_text <- variable_textdriver
      } else {
        variable_text <- "" 
           }
      
      text_data <- data.frame(category = c("Early", "Undetermined", "Late"),
                              text = c(early_text, variable_text, late_text),
                              x = c(0.165, 0.495, 0.83),
                              y = c(0.5, 0.5, 0.5),
                              hjust = c(0.5, 0.5, 0.5),
                              vjust = c(0.5, 0.5, 0.5))
      
      title_data <- data.frame(x = c(0.165, 0.495, 0.83),
                               y = rep(0.95, 3),
                               label = c("Early", "Undetermined", "Late"))
      
      bg_data <- data.frame(
        xmin = c(0, 0.33, 0.66),
        xmax = c(0.33, 0.66, 1),
        ymin = c(0, 0, 0),
        ymax = c(1, 1, 1),
        fill = color_summary)
      
      # Create the base plot with colored background sections
      base_plot <- ggplot() +
        geom_rect(data = bg_data, aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = fill), alpha = 0.9) +
        scale_fill_identity() +
        theme_void()
      
      for (i in 1:nrow(text_data)) {
        base_plot <- base_plot +
          annotate("text", x = text_data$x[i], y = text_data$y[i], label = text_data$text[i],
                   hjust = text_data$hjust[i], vjust = text_data$vjust[i], size = 6)
        
      }
      
      for (i in 1:nrow(title_data)) {
        base_plot <- base_plot +
          annotate("text", x = title_data$x[i], y = title_data$y[i], label = title_data$label[i],
                   hjust = 0.5, vjust = 1, size = 8, fontface = "bold")
      }
      
      ggsave(filename = paste0(path_timeline, "/Summary_", type, ".pdf"), plot = base_plot, width = 12, height = 5, units = "in")
      
      plot_com2 <- plot_barcom + base_plot + patchwork::plot_layout(widths = c(3, 4)) +
        plot_annotation(
          title = paste0("Timeline of ", type),
          theme = theme(plot.title = element_text(size = 24, face = "bold", hjust = 0.5)))
      
      ggsave(filename = paste0(path_timeline, "/Timelinebar_", type, ".pdf"), plot = plot_com2,  width = 28, height = 7, units = "in")
      
      rm(Freq)
      rm(data_bar)
      rm(plot_com1)
      rm(plot_group)
      rm(plot_com2)
      rm(plot_bar)
    }
  }
}
```

## CRC driver timing
```{r}
# "6ca5c1bb-275b-4d05-948a-3c6c7d03fab9" was not included in Toby's results
type = "ColoRect-AdenoCA"
figure_path <- ".../Project/MHC_evolution/GRITIC/output/250717/summary250717/timeline/"
gritic_path = ".../Project/MHC_evolution/GRITIC/output/250717/summary250717/"
path_timeline = ".../Project/MHC_evolution/GRITIC/output/250717/summary250717/timeline/"

genelist = genelist_APC_CRC <- unique(c("APC", "FBXW7", "KRAS", "TP53", "CTNNB1", "SOX9", "SOX9", "GNAS", "TCF7L2", "SMAD4", "FAT4"))
diff_all <- fread(paste0(gritic_path, "diff_alldrivergenes.csv")) %>% filter(aliquot_id %in% sample_MSS) %>% filter(aliquot_id %in% sample_sucess)

data_all <- diff_all %>% filter(histology_abbreviation == "ColoRect-AdenoCA", regulator == "Driver", pathway %in% genelist)
  
data_filter1 <- data_all %>% group_by(pathway, histology_abbreviation) %>%
    summarise(count = n()) %>%
    ungroup() %>%
    inner_join(., data_all, by = c("histology_abbreviation", "pathway")) %>%
    mutate(n_base = 1,
             percent_base = n_base/count,
             n_late = 250 * late_ratio,
             n_undetermined = 250 * undetemined_ratio,
             ratio_plot = (n_late + n_undetermined/2) / 250,
             timing_bin = cut(ratio_plot, breaks = range_breaks, include.lowest = TRUE)) %>% 
    arrange(desc(count))
  
data_filter <- data_filter1 %>% filter(timing_cat != "Undetermined") %>%
    group_by(pathway) %>% 
    summarise(Sum_determined = n()) %>% 
    ungroup %>% 
    inner_join(data_filter1, ., by = "pathway") 
  
Freq <- as.data.frame(table(data_filter$pathway, data_filter$timing_cat))
Freq <- Freq %>% setNames(c("pathway", "time_period", "Freq")) 
type_sum <- as.data.frame(table(LOH_MSI$histology_abbreviation)) %>% filter(Var1 == !!type) %>% pull(Freq)
    
Sum <- as.data.frame(table(data_filter$pathway)) %>% 
mutate(Type_sum = type_sum) %>% 
setNames(c("pathway", "Sum", "Sum_type")) %>% 
mutate(Prevelance = Sum/Sum_type) 
    
data_prop <- inner_join(Freq, Sum, by = "pathway") %>% mutate(Proportion = Freq/Sum)

ratio <- data_prop %>% filter(time_period != "Undetermined") %>%
        group_by(pathway, time_period) %>% 
        summarise(sum_freq = sum(Freq), .groups = 'drop') %>% 
        pivot_wider(names_from = time_period, values_from = sum_freq) %>% 
        mutate(ratio = (Late + 1)/(Early + 1)) %>% 
        inner_join(., data_prop, by = "pathway") %>%
        distinct(pathway, .keep_all = TRUE) %>% 
        mutate(pro_early = Early/Sum,
               pro_late = Late/Sum) %>% 
        arrange(desc(ratio)) 
      
order_all <- ratio %>% arrange(desc(ratio)) %>% pull(pathway)
      
data_bar <- data_filter %>%
        mutate(pathway = factor(pathway, levels = order_all),
               timing_bin = factor(timing_bin, levels = rev(timing_bin_levels)))
      
timing_propCRC <- ggplot(data_bar, aes(x = pathway, fill = timing_bin)) +
        geom_bar(position = "fill") +
        theme_minimal() +
        labs(x = "",  y = "Proportion", fill = "Timing Probability", title = "") +
        scale_fill_manual(values = color_vector, drop = FALSE) +
        theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.5, size = 18, colour = "black"),
              axis.text.y = element_text(size = 16, colour = "black", hjust = 1),
              plot.title = element_text(size = 20, vjust = 0.5),
              axis.title.y = element_text(size = 18),
              axis.title.x = element_text(size = 18),
              legend.text = element_blank(),  
              legend.title = element_text(size = 16),
              legend.position = "none",
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              plot.margin = margin(0, 30, 5, 5)) +
        coord_flip()

plot_CRC <- cowplot::plot_grid(timing_propCRC, 
      legend, 
      ncol = 2,  
      rel_widths = c(1.5, 0.3))
   
ggsave(filename = paste0(figure_path, "/Figure2a_ColoRect-AdenoCA.pdf"), plot = plot_CRC,  width = 5, height = 6, units = "in")
ggsave(filename = paste0(path_timeline, "/Figure2a_ColoRect-AdenoCA.pdf"), plot = plot_CRC,  width = 5, height = 6, units = "in")
```

## DLBCL APM timing
```{r}
type = "ColoRect-AdenoCA"
figure_path <- ".../Project/MHC_evolution/GRITIC/output/250717/summary250717/timeline/"
genelist = genelist_APC_CRC <- unique(c("APC", "FBXW7", "KRAS", "TP53", "CTNNB1", "SOX9", "SOX9", "GNAS", "TCF7L2", "SMAD4", "FAT4"))
diff_all <- fread(paste0(gritic_path, "diff_alldrivergenes.csv")) %>% filter(aliquot_id %in% sample_MSS)
data_all <- diff_all %>% filter(histology_abbreviation == "ColoRect-AdenoCA", regulator == "Driver", pathway %in% genelist)
  
data_filter1 <- data_all %>% group_by(pathway, histology_abbreviation) %>%
    summarise(count = n()) %>%
    ungroup() %>%
    inner_join(., data_all, by = c("histology_abbreviation", "pathway")) %>%
    mutate(n_base = 1,
             percent_base = n_base/count,
             n_late = 250 * late_ratio,
             n_undetermined = 250 * undetemined_ratio,
             ratio_plot = (n_late + n_undetermined/2) / 250,
             timing_bin = cut(ratio_plot, breaks = range_breaks, include.lowest = TRUE)) %>% 
    arrange(desc(count))
  
data_filter <- data_filter1 %>% filter(timing_cat != "Undetermined") %>%
    group_by(pathway) %>% 
    summarise(Sum_determined = n()) %>% 
    ungroup %>% 
    inner_join(data_filter1, ., by = "pathway") 
  
Freq <- as.data.frame(table(data_filter$pathway, data_filter$timing_cat))
Freq <- Freq %>% setNames(c("pathway", "time_period", "Freq")) 
type_sum <- as.data.frame(table(LOH_MSI$histology_abbreviation)) %>% filter(Var1 == !!type) %>% pull(Freq)
    
Sum <- as.data.frame(table(data_filter$pathway)) %>% 
mutate(Type_sum = type_sum) %>% 
setNames(c("pathway", "Sum", "Sum_type")) %>% 
mutate(Prevelance = Sum/Sum_type) 
    
data_prop <- inner_join(Freq, Sum, by = "pathway") %>% mutate(Proportion = Freq/Sum)

ratio <- data_prop %>% filter(time_period != "Undetermined") %>%
        group_by(pathway, time_period) %>% 
        summarise(sum_freq = sum(Freq), .groups = 'drop') %>% 
        pivot_wider(names_from = time_period, values_from = sum_freq) %>% 
        mutate(ratio = (Late + 1)/(Early + 1)) %>% 
        inner_join(., data_prop, by = "pathway") %>%
        distinct(pathway, .keep_all = TRUE) %>% 
        mutate(pro_early = Early/Sum,
               pro_late = Late/Sum) %>% 
        arrange(desc(ratio)) 
      
order_all <- ratio %>% arrange(desc(ratio)) %>% pull(pathway)
      
data_bar <- data_filter %>%
        mutate(pathway = factor(pathway, levels = order_all),
               timing_bin = factor(timing_bin, levels = rev(timing_bin_levels)))
      
timing_propCRC <- ggplot(data_bar, aes(x = pathway, fill = timing_bin)) +
        geom_bar(position = "fill") +
        theme_minimal() +
        labs(x = "",  y = "Proportion", fill = "Timing Probability", title = "") +
        scale_fill_manual(values = color_vector, drop = FALSE) +
        theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.5, size = 18, colour = "black"),
              axis.text.y = element_text(size = 16, colour = "black", hjust = 1),
              plot.title = element_text(size = 20, vjust = 0.5),
              axis.title.y = element_text(size = 18),
              axis.title.x = element_text(size = 18),
              legend.text = element_blank(),  
              legend.title = element_text(size = 16),
              legend.position = "none",
              panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              plot.margin = margin(0, 30, 5, 5)) +
        coord_flip()

plot_CRC <- cowplot::plot_grid(timing_propCRC, 
      legend, 
      ncol = 2,  
      rel_widths = c(1.5, 0.3))
   
ggsave(filename = paste0(figure_path, "/Figure2a_ColoRect-AdenoCA.pdf"), plot = plot_CRC,  width = 5, height = 6, units = "in")
ggsave(filename = paste0(path_timeline, "/Figure2a_ColoRect-AdenoCA.pdf"), plot = plot_CRC,  width = 5, height = 6, units = "in")

type = "Lymph-BNHL"
figure_path <- ".../Project/MHC_evolution/GRITIC/output/250717/summary250717/timeline/"
figure_path <- ".../Project/Figures/20250808/Figure2E/"
APM_timing <- fread(".../Project/MHC_evolution/GRITIC/output/250717/summary250717/timing_APM.csv") %>% 
  filter(histology_abbreviation == "Lymph-BNHL") 
sample_APM_timingBNHL <- APM_timing %>% pull(aliquot_id) %>% unique()

df_background <- fread(paste0(gritic_path, "/timing_driverbackground.csv")) %>% 
  filter(!is.na(mean_background)) %>%
  mutate(mean_timing = mean_background,
         median_timing = median_background,
         pathway = event,
         regulator = "backgound") %>%
  dplyr::select(c("sample_id", "Hugo_Symbol", "event", "Other_Info", everything()), -mean_background, -median_background)

df_columns <- names(df_background)

df_driver <- fread(paste0(gritic_path, "/timing_allmutations.csv")) %>% mutate(Hugo_Symbol = event) %>%
  dplyr::select(c("sample_id", "Hugo_Symbol", "event", "Other_Info", everything())) %>%
  dplyr::select(all_of(df_columns)) %>% 
  filter(sample_id %in% sample_APM_timingBNHL) 

sample_background <- unique(df_background$sample_id)
sample_mut <- unique(df_driver$sample_id)
sample_comm <- intersect(sample_background, sample_mut)

df_final <- rbind(df_background, df_driver) %>% filter(sample_id %in% sample_comm) %>%
  distinct()

## Compare the genes with background one by one
diff_driver <- list()

sample_list <- unique(df_final$sample_id)

for (sample in sample_list) {
  
  subset <- df_final %>% filter(sample_id %in% sample) %>% dplyr::select(sample_id, Hugo_Symbol, X0:X249)
  genes <- unique(subset$Hugo_Symbol)
  genes <- genes[genes != "Background"]
  
  if(length(genes) > 0) {
    
    for (gene in genes) {
      
      gene1_data <- subset %>% filter(Hugo_Symbol == !!gene) %>% dplyr::select(X0:X249)
      gene2_data <- subset %>% filter(Hugo_Symbol == "Background") %>% dplyr::select(X0:X249)
      
      diff_values <- gene1_data - gene2_data
      
      pair <- paste0(sample, "_", gene)
      
      print(pair)
      
      diff_driver[[pair]] <- cbind(sample_id = sample, event1 = gene, diff_values)
      
    }
  }
}

diff_drivergene <- rbindlist(diff_driver, fill = TRUE) 
write.csv(diff_drivergene, file = paste0(gritic_path, "/timing_pair_APM_timingBNHL.csv"), row.names = F)

diff_drivergene <- fread(paste0(gritic_path, "/timing_pair_APM_timingBNHL.csv")) %>% 
  mutate(Hugo_Symbol = sub("_.*", "", event1)) %>%
  group_by(sample_id, Hugo_Symbol) %>%
  summarise(across(all_of(column_names), ~ mean(., na.rm = TRUE)), .groups = "drop") %>%
  mutate(pathway = Hugo_Symbol,
         early_ratio = rowMeans(dplyr::select(., all_of(column_names)) < 0, na.rm = TRUE),
         late_ratio = rowMeans(dplyr::select(., all_of(column_names)) > 0, na.rm = TRUE),
         undetemined_ratio = rowMeans(dplyr::select(., all_of(column_names)) == 0, na.rm = TRUE),
         timing_cat = case_when(early_ratio > cutoff_mut ~ "Early", 
                                late_ratio > cutoff_mut ~ "Late", 
                                TRUE ~ "Undetermined"),
         aliquot_id = sample_id,
         histology_abbreviation = "Lymph-BNHL") %>% 
  filter(Hugo_Symbol %in% !!APM) %>%
  dplyr::select(c("sample_id", "aliquot_id", "histology_abbreviation", "Hugo_Symbol", "pathway", everything())) 
write_csv(diff_drivergene, paste0(gritic_path, "/diff_allmutations_APM_timingBNHL.csv"))

data_all <- fread(paste0(gritic_path, "/diff_allmutations_APM_timingBNHL.csv"))
data_filter1 <- data_all %>% 
  group_by(pathway, histology_abbreviation) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  inner_join(., data_all, by = c("histology_abbreviation", "pathway")) %>%
  mutate(n_base = 1,
             percent_base = n_base/count,
             n_late = 250 * late_ratio,
             n_undetermined = 250 * undetemined_ratio,
             ratio_plot = (n_late + n_undetermined/2) / 250,
             timing_bin = cut(ratio_plot, breaks = range_breaks, include.lowest = TRUE)) %>% 
   arrange(desc(count))

data_filter <- data_filter1 %>% #filter(timing_cat != "Undetermined") %>%
  group_by(pathway) %>% 
  summarise(Sum_determined = n()) %>% 
  ungroup %>% 
  inner_join(data_filter1, ., by = "pathway") 

Freq <- as.data.frame(table(data_filter$pathway, data_filter$timing_cat))
Freq <- Freq %>% setNames(c("pathway", "time_period", "Freq")) 
type_sum <- as.data.frame(table(LOH_MSI$histology_abbreviation)) %>% filter(Var1 == !!type) %>% pull(Freq)

Sum <- as.data.frame(table(data_filter$pathway)) %>% 
  mutate(Type_sum = type_sum) %>% 
  setNames(c("pathway", "Sum", "Sum_type")) %>% 
  mutate(Prevelance = Sum/Sum_type) 

data_prop <- inner_join(Freq, Sum, by = "pathway") %>% mutate(Proportion = Freq/Sum)

ratio <- data_prop %>% filter(time_period != "Undetermined") %>%
  group_by(pathway, time_period) %>% 
  summarise(sum_freq = sum(Freq), .groups = 'drop') %>% 
  pivot_wider(names_from = time_period, values_from = sum_freq) %>% 
  mutate(ratio = (Late + 1)/(Early + 1)) %>% 
  inner_join(., data_prop, by = "pathway") %>%
  distinct(pathway, .keep_all = TRUE) %>% 
  mutate(pro_early = Early/Sum,
         pro_late = Late/Sum) %>% 
  arrange(desc(ratio)) 

order_all <- ratio %>% arrange(desc(ratio)) %>% pull(pathway)

data_bar <- data_filter %>%
  mutate(pathway = factor(pathway, levels = order_all),
         timing_bin = factor(timing_bin, levels = rev(timing_bin_levels)))

timing_propCRC <- ggplot(data_bar, aes(x = pathway, fill = timing_bin)) +
  geom_bar(position = "fill") +
  theme_minimal() +
  labs(x = "",  y = "Proportion", fill = "Timing Probability", title = "") +
  scale_fill_manual(values = color_vector, drop = FALSE) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.5, size = 18, colour = "black"),
        axis.text.y = element_text(size = 16, colour = "black", hjust = 1),
        plot.title = element_text(size = 20, vjust = 0.5),
        axis.title.y = element_text(size = 18),
        axis.title.x = element_text(size = 18),
        legend.text = element_blank(),  
        legend.title = element_text(size = 16),
        legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(0, 30, 5, 5)) +
  coord_flip()

plot_CRC <- cowplot::plot_grid(timing_propCRC, 
                               legend, 
                               ncol = 2,  
                               rel_widths = c(1.5, 0.3))
ggsave(filename = paste0(figure_path, "/Figure2e_APMpathway.pdf"), plot = plot_CRC,  width = 5, height = 6, units = "in")
```

# APM timing
```{r}
mode = 2
diff_all <- fread(paste0(path_timeline, "/diff_all.csv"))

data_all <- diff_all %>% filter(pathway == "APM") %>% 
  filter(aliquot_id %in% sample_MSS) %>% 
  filter(aliquot_id %in% sample_sucess)

data_filter1 <- data_all %>% group_by(pathway, histology_abbreviation) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  inner_join(., data_all, by = c("histology_abbreviation", "pathway")) %>%
  mutate(n_base = 1,
             percent_base = n_base/count,
             n_late = 250 * late_ratio,
             n_undetermined = 250 * undetemined_ratio,
             ratio_plot = (n_late + n_undetermined/2) / 250,
             timing_bin = cut(ratio_plot, breaks = range_breaks, include.lowest = TRUE)) %>% 
      arrange(desc(count))%>%
  filter(count > 4)

data_filter <- data_filter1 %>% #filter(timing_cat != "Undetermined") %>%
  group_by(histology_abbreviation) %>% 
  summarise(Sum_determined = n()) %>% 
  ungroup %>% 
  inner_join(data_filter1, ., by = "histology_abbreviation")  
Freq <- as.data.frame(table(data_filter$histology_abbreviation, data_filter$timing_cat))
Freq <- Freq %>% setNames(c("histology_abbreviation", "time_period", "Freq")) 

ratio <- Freq %>% filter(time_period != "Undetermined") %>%
  group_by(histology_abbreviation, time_period) %>% 
  summarise(sum_freq = sum(Freq), .groups = 'drop') %>% 
  pivot_wider(names_from = time_period, values_from = sum_freq) %>% 
  mutate(ratio = (Late + 1)/(Early + 1),
         histology_abbreviation = as.character(histology_abbreviation)) %>% 
  arrange(desc(ratio)) 

order_all <- ratio %>% arrange(desc(ratio)) %>% pull(histology_abbreviation)

data_bar <- data_filter %>% 
  #filter(!histology_abbreviation %in% c("Lung−AdenoCA", "Cervix−SCC", "Breast−LobularCA", "Kidney−ChRCC", "CNS−Oligo", "CNS−Medullo", "CNS−GBM",
  #                                      "Thy−AdenoCA", "Prost−AdenoCA", "Panc−Endocrine", "Myeloid−MPN")) %>% 
  mutate(histology_abbreviation = factor(histology_abbreviation, levels = order_all),
         timing_bin = factor(timing_bin, levels = rev(timing_bin_levels))) 

timing_propAPM <- ggplot(data_bar, aes(x = histology_abbreviation, fill = timing_bin)) +
  geom_bar(position = "fill") +
  theme_minimal() +
  labs(x = "",  y = "Proportion", fill = "Timing Probability", title = "") +
  scale_fill_manual(values = color_vector, drop = FALSE) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.5, size = 18, colour = "black"),
        axis.text.y = element_text(size = 16, colour = "black", hjust = 1),
        plot.title = element_text(size = 20, vjust = 0.5),
        axis.title.y = element_text(size = 18),
        axis.title.x = element_text(size = 18),
        legend.text = element_blank(),  
        legend.title = element_text(size = 16),
        legend.position = "none",
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        plot.margin = margin(0, 30, 5, 5)) +
  coord_flip()

plot_APM <- cowplot::plot_grid(timing_propAPM, 
      legend, 
      ncol = 2,  
      rel_widths = c(1.5, 0.3))

ggsave(filename = paste0(path_timeline, "/Figure2c_APM.pdf"), plot = plot_APM,  width = 6, height = 6, units = "in")
ggsave(filename = "/Users/wchen20/Desktop/Figure2c_APM.pdf", plot = plot_APM,  width = 6, height = 6, units = "in")
```
