---
title: "R1Major5_uncertainty_MedianTiming_250samplingFinal.Rmd"
author: "AJ"
date: "2026-06-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load packacges
```{r}
rm(list = ls())

suppressWarnings(library("stringr"))
suppressWarnings(library("dplyr"))
suppressWarnings(library("patchwork"))
suppressWarnings(library("ggpubr"))
suppressWarnings(library("ggridges"))
suppressWarnings(library("truncnorm"))
suppressWarnings(library("tidyr"))
suppressWarnings(library("data.table"))
suppressWarnings(library("readr"))
suppressWarnings(library("ggplot2"))
suppressWarnings(library("ggrepel"))
suppressWarnings(library("tibble"))
suppressWarnings(library("textshape"))
suppressWarnings(library("matrixStats"))
suppressWarnings(library("bayestestR"))
suppressWarnings(library("BradleyTerry2"))
suppressWarnings(library("purrr"))
suppressWarnings(library("tidyr"))
suppressWarnings(library("purrr"))
suppressWarnings(library("cowplot"))
suppressWarnings(library("ggsci"))
suppressWarnings(library("randomcoloR"))
suppressWarnings(library("PlackettLuce"))

## Path
PCAWG_path <- "/rsrch6/scratch/hema_bio-Malignan/wchen20/PCAWG"
PCAWG_path <- "/Users/wchen20/Desktop/PCAWG"
gritic_path <- paste0(PCAWG_path, "/MHC_evolution/GRITIC/output")

source("/Users/wchen20/Desktop/PCAWG/code/source/basic.R")
source("/Users/wchen20/Desktop/PCAWG/code/source/insert_newlines.R")
source("/Users/wchen20/Desktop/PCAWG/code/source/plot_segTiming.R")

## Timing of WGD
wgd_median <- fread("/Users/wchen20/Desktop/PCAWG/MHC_evolution/GRITIC/output/wgd_median.csv") %>% 
  dplyr::select(wgd_mediantiming, sample_id) %>% 
  filter(sample_id %in% LOH_MSI$aliquot_id)

sample_wgd <- wgd_median %>% pull(sample_id)

## Save Path
#gritic_path = "/Users/wchen20/Desktop/PCAWG/MHC_evolution/GRITIC/output/250717/summary250717/"
#path_timeline = "/Users/wchen20/Desktop/PCAWG/MHC_evolution/GRITIC/output/250717/summary250717/timeline/"

sample_HLA <- fread("/Users/wchen20/Desktop/PCAWG/MHC_evolution/GRITIC/output/250717/summary250717/sample_HLA.csv") %>% pull(sample_id)
```

# 0. Set path
```{r}
path_gritic = "/Users/wchen20/Desktop/PCAWG/MHC_evolution/GRITIC/output/allmutation/mafgritic_nonsyn/"
path_gritic_pathway = "/Users/wchen20/Desktop/PCAWG/MHC_evolution/GRITIC/output/allmutation/mafgritic_nonsyn/pathway/min_drivermutation_250/"

path_HLALOH = "/Users/wchen20/Desktop/2026-02-16_Draft/Final_Results/1_APM/2_HLALOH/"
```

# 1. Timing for nonsyn
```{r}
nonsyn_filter_nonPloy <- fread(paste0(path_gritic, "/GRITIC_success.csv"))
sample_sucess <- unique(nonsyn_filter_nonPloy$sample_id)

mafgritic_files <- list.files(path = path_gritic, pattern = "mafgritic_nonsyn_*", recursive = FALSE, full.names = TRUE)
data_timing_driver <- list()

for (file in mafgritic_files) {
  
  type = basename(file)
  type = sub("mafgritic_nonsyn_", "", type)
  type = sub(".csv", "", type)
  
  data_timing_driver[[type]] <- fread(file) %>% 
    filter(Variant_Classification %in% non_syn)
  
}

data_timing_driver_all <- rbindlist(data_timing_driver) %>% 
  filter(Tumor_Sample_Barcode %in% sample_MSS) %>%
  filter(Tumor_Sample_Barcode %in% sample_sucess) %>%
  mutate(median_value = rowMedians(as.matrix(dplyr::select(., all_of(column_names))), na.rm = TRUE),
         early_ratio = rowMeans(dplyr::select(., all_of(column_names)) < 0.5, na.rm = TRUE)) 

write.csv(data_timing_driver_all, file = paste0(path_gritic_pathway, "/nonsyn_filter_diff.csv"), row.names = F)

sample_all <- unique(data_timing_driver_all$Tumor_Sample_Barcode)
missing_ids <- setdiff(sample_all, sample_sucess)

check <- LOH_MSI %>% filter(aliquot_id %in% missing_ids)
```

# 2. Timing for driver genes
```{r}
drivergene_cancer <- driver_list %>% pull(Gene)

data_timing_driver_all <- fread(paste0(path_gritic_pathway, "/nonsyn_filter_diff.csv")) %>% 
  filter(Hugo_Symbol %in% drivergene_cancer) %>%
  mutate(sample_id = Tumor_Sample_Barcode,
         pathway = Hugo_Symbol,
         event = Hugo_Symbol,
         pathway_new = Hugo_Symbol) %>%
  dplyr::select(c("sample_id", "Hugo_Symbol", "event", everything()))
write.csv(data_timing_driver_all, file = paste0(path_gritic_pathway, "/diff_driver.csv"), row.names = F)

length(unique(data_timing_driver_all$sample_id))
kept_cols <- names(data_timing_driver_all)

data_timing_driver_all <- fread(paste0(path_gritic_pathway, "/nonsyn_filter_diff.csv")) %>% 
  filter(Hugo_Symbol %in% drivergene_cancer) %>%
  group_by(Tumor_Sample_Barcode, Hugo_Symbol) %>%
  slice_min(median_value, n = 1, with_ties = FALSE) %>%
  mutate(sample_id = Tumor_Sample_Barcode,
         pathway = Hugo_Symbol,
         event = Hugo_Symbol,
         pathway_new = Hugo_Symbol) %>%
  dplyr::select(c("sample_id", "Hugo_Symbol", "event", everything()))
write.csv(data_timing_driver_all, file = paste0(path_gritic_pathway, "/diff_driver.csv"), row.names = F)
```

# 3. Timing for driver mutations
```{r}
mafgritic_files <- list.files(path = path_gritic, pattern = "mafgritic_drivermutation_*", recursive = FALSE, full.names = TRUE)
data_timing_driver <- list()

for (file in mafgritic_files) {
  
  type = basename(file)
  type = sub("mafgritic_drivermutation_", "", type)
  type = sub(".csv", "", type)
  
  data_timing_driver[[type]] <- fread(file)
  
}

data_timing_drivermutations <- rbindlist(data_timing_driver) %>% 
    filter(Tumor_Sample_Barcode %in% sample_MSS) %>%
    filter(Tumor_Sample_Barcode %in% sample_sucess) %>%
    mutate(sample_id = Tumor_Sample_Barcode,
         pathway = Hugo_Symbol,
         event = Hugo_Symbol,
         pathway_new = Hugo_Symbol,
         median_value = rowMedians(as.matrix(across(all_of(column_names))), na.rm = TRUE),
         early_ratio = rowMeans(dplyr::select(., all_of(column_names)) < 0.5, na.rm = TRUE)) %>% 
  group_by(sample_id, pathway) %>%
  slice_min(median_value, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  dplyr::select(all_of(kept_cols))
  
write.csv(data_timing_drivermutations, file = paste0(path_gritic_pathway, "/diff_drivermutations.csv"), row.names = F)
length(unique(data_timing_drivermutations$sample_id))

table(data_timing_drivermutations$category)
Freq <- as.data.frame(table(data_timing_drivermutations$Hugo_Symbol, data_timing_drivermutations$histology_abbreviation))
```

# 4. Timing for APM
```{r}
data_timing_APM_all <- fread(paste0(path_gritic_pathway, "/nonsyn_filter_diff.csv"))  %>% 
  filter(Tumor_Sample_Barcode %in% sample_MSS) %>%
  filter(Tumor_Sample_Barcode %in% sample_sucess) %>%
  filter(Hugo_Symbol %in% APM) %>%
  mutate(subclonal_ratio = rowMeans(dplyr::select(., all_of(column_names)) == 1.01, na.rm = TRUE),
         subclonal = ifelse(subclonal_ratio > 0.5, "subclonal", "clonal")) %>%
  group_by(Tumor_Sample_Barcode) %>%
  #slice_min(abs(median_value - median(median_value, na.rm = TRUE)), n = 1, with_ties = FALSE) %>% # Median
  slice_min(median_value, n = 1, with_ties = FALSE) %>% #minimun
  #slice_min(hdi_width, n = 1, with_ties = FALSE) %>%
  #slice_max(early_ratio, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  mutate(
    sample_id = Tumor_Sample_Barcode,
    Hugo_Symbol = "APM",
    pathway = "APM",
    event = "APM",
    pathway_new = "APM"
  ) %>%
  dplyr::select(all_of(kept_cols))

write.csv(data_timing_APM_all, file = paste0(path_gritic_pathway, "/diff_APM.csv"), row.names = F)
```

# 5. Timing for pathways
## Each pathway
```{r}
data_timing_ratio <- fread(paste0(path_gritic_pathway, "/nonsyn_filter_diff.csv"))

cell_types <- c("MHCregulators", "Tcells", "NKcells", "Macrophages", "GammadeltaTcells")
regulators <- c("pos", "neg")

for (cell_type in cell_types) {
  
   CRISPR_path <- paste0("/Users/wchen20/Desktop/2026-02-16_Draft/Revision/R1Major2/005/")
   
   for (reg in regulators) {
    
      CRISPR_top10 <- fread(paste0(CRISPR_path, "/Summary/", cell_type, "_C5BP_", reg, "_list.csv"), header = T) %>% 
        dplyr::mutate(Pathway = ID, Pathway1 = Description) %>%
        dplyr::select(Pathway, Pathway1)
      
      CRISPR_pathways <- fread(paste0(CRISPR_path, "/Summary/", cell_type, "_C5BP_", reg, "_allgenelist.csv")) %>% 
        inner_join(CRISPR_top10, by = c("Pathway", "Pathway1")) %>%
        dplyr::select(Gene, Pathway1)
     
     pathway_list <- unique(CRISPR_pathways$Pathway1)
        
     for (pathway in pathway_list) {
       
        pathway_title <- str_to_title(pathway)
        pathway <- sub("/", "-", pathway)
        plot_key <- paste0(cell_type, "_", reg, "_", pathway)
        
        CRISPR_gene <- CRISPR_pathways %>% filter(Pathway1 == !!pathway) %>% pull(Gene)
          
        tryCatch({
          
          df_Reg <- data_timing_ratio %>%
             filter(Hugo_Symbol %in% !!CRISPR_gene) %>%
             filter(Variant_Classification %in% non_syn) %>%
              group_by(Tumor_Sample_Barcode) %>%
              #slice_min(abs(median_value - median(median_value, na.rm = TRUE)), n = 1, with_ties = FALSE) %>%
              slice_min(median_value, n = 1, with_ties = FALSE) %>% 
              #slice_min(hdi_width, n = 1, with_ties = FALSE) %>%
              #slice_max(early_ratio, n = 1, with_ties = FALSE) %>%
              ungroup() %>% 
              mutate(sample_id = Tumor_Sample_Barcode,
                    Hugo_Symbol = pathway,
                    event = pathway) %>%
             dplyr::select(c("sample_id", "Hugo_Symbol", "event", everything()))
 
           write.csv(df_Reg, file = paste0(path_gritic_pathway, "/difference_", cell_type, "_", reg, "_", pathway, ".csv"), row.names = F)
           
             }, error = function(e) {
                    message("Error processing ", cell_type, "_", reg, "_", pathway, ": ", e$message)
             })
     }
   }
}
```

## Combined
```{r}
rename_list <- fread("/Users/wchen20/Desktop/PCAWG/CRISPR/CRISPRscreen20240118/rename_list.csv") %>% 
  mutate(pathway = str_to_title(pathway),
         pathway = sub("/", "-", pathway))

file_paths <- list.files(path = path_gritic_pathway, pattern = "difference_*", recursive = FALSE, full.names = TRUE)
data_list <- list()

for (file in file_paths) {
       
     file_name <- basename(file)
     file_name <- sub("difference_", "", file_name)
     file_name <- sub(".csv", "", file_name)
     data_list[[file_name]] <- fread(file) %>% mutate(Study = file_name)
}

diff_pathway <- tidy_list(data_list) %>% 
  tidyr::separate(Study, into = c("celltype", "regulator", "pathway"), extra = "merge", fill = "right")  %>%
  mutate(pathway = str_to_title(pathway),
         aliquot_id = Tumor_Sample_Barcode) %>% 
  inner_join(rename_list,., by = "pathway") %>% 
  mutate(pathway_new = case_when(pathway_new == "Antigen Presentation via MHC-I" & celltype == "MHCregulators" ~ paste0(pathway_new, " (MHC-I) "),
                                 pathway_new == "Antigen Presentation via MHC-I" & celltype == "Tcells" ~ paste0(pathway_new, " (Tcells) "),
                                 pathway_new == "Antigen Processing and Presentation" & celltype == "NKcells" ~ paste0(pathway_new, " (NKcells) "),
                                 pathway_new == "IFNg Signaling" & celltype == "NKcells" ~ paste0(pathway_new, " (NKcells) "),
                                 pathway_new == "IFNg Signaling" & celltype == "MHCregulators" ~ paste0(pathway_new, " (MHC-I)"),
                                 pathway_new == "IFNg Signaling" & celltype == "Tcells" ~ paste0(pathway_new, " (Tcells)"),
                                 pathway_new == "Necroptosis" & celltype == "Tcells" ~ paste0(pathway_new, " (Tcells)"),
                                 pathway_new == "Necroptosis" & celltype == "NKcells" ~ paste0(pathway_new, " (NKcells)"),
                                 pathway_new == "Regulation of Autophagy" & celltype == "MHCregulators" ~ paste0(pathway_new, " (MHC-I)"),
                                 pathway_new == "Autophagy" & celltype == "Tcells" ~ paste0(pathway_new, " (Tcells)"),
                                 pathway_new == "TNF Pathway" & celltype == "Tcells" ~ paste0(pathway_new, " (Tcells)"),
                                 pathway_new == "TNF Pathway" & celltype == "NKcells" ~ paste0(pathway_new, " (NKcells)"),
                                 pathway_new == "IFNa Pathway" & celltype == "MHCregulators" ~ paste0(pathway_new, " (MHC-I)"),
                                 pathway_new == "IFNa Pathway" & celltype == "GammadeltaTcells" ~ paste0(pathway_new, " (gdTcells)"),
                                 pathway_new == "IFNg Pathway" & celltype == "GammadeltaTcells" ~ paste0(pathway_new, " (gdTcells)"),
                                 TRUE ~ pathway_new)) %>%
  mutate(pathway = pathway_new) %>%
  dplyr::select(all_of(kept_cols), "celltype", "regulator")

write.csv(diff_pathway, file = paste0(path_gritic_pathway, "/diff_all.csv"), row.names = F)
```

# 6. Timing for copy number losses
## Timing of losses
```{r}
INFy_loss <- fread("/Users/wchen20/Desktop/2026-02-16_Draft/Revision/R2Major9/IFNy_loss.csv") %>%
  distinct(sample_id, .keep_all = F) %>% 
  mutate(event = "IFNgLoss")

HLALOH_path <- "/Users/wchen20/Desktop/PCAWG/MHC_evolution/HLALOH/results/"
df_HLALOH <- fread(paste0(HLALOH_path, "/df_HLALOH.csv")) %>%
  distinct(sample_id, .keep_all = F)  %>% 
  mutate(event = "HLALOH")

df_loss <- df_HLALOH %>%
  rbind(INFy_loss, .) %>%
  filter(sample_id %in% sample_wgd) %>%
  filter(sample_id %in% sample_sucess) %>%
  filter(sample_id %in% sample_MSS) %>%
  inner_join(., wgd_median, by = "sample_id") %>%
  rowwise() %>%
  mutate(new_data = list(map_dfc(1:250, ~ runif(1, min = 0, max = wgd_mediantiming - epsilon)))) %>%
  unnest_wider(new_data) %>%
  rename_with(~ paste0("X", 0:249), .cols = starts_with("...")) %>% 
  ungroup() %>% 
  mutate(
         median_value = rowMedians(as.matrix(dplyr::select(., all_of(column_names))), na.rm = TRUE),
         early_ratio = rowMeans(dplyr::select(., all_of(column_names)) < 0.5, na.rm = TRUE),
         pathway = event,
         pathway_new = event,
         Tumor_Sample_Barcode = sample_id,
         Hugo_Symbol = event,
         Chromosome = NA, 
         Start_Position = NA, 
         End_Position= NA, 
         Reference_Allele = NA, 
         Tumor_Seq_Allele1 = NA,
         Tumor_Seq_Allele2 = NA,
         Variant_Classification = "Missense_Mutation", 
         Variant_Type = "CNV") %>% 
  inner_join(LOH_type, by = "sample_id")
  
write.csv(df_loss, file = paste0(path_gritic_pathway, "/diff_loss.csv"), row.names = F)
```

# 9. Timing for WGD
```{r}
wgd_median <- fread("/Users/wchen20/Desktop/PCAWG/MHC_evolution/GRITIC/output/wgd_median.csv") %>%
        filter(Tumor_Sample_Barcode %in% sample_sucess) %>%
        filter(Tumor_Sample_Barcode %in% sample_MSS) %>%
        mutate(pathway = "WGD",
               median_value = wgd_mediantiming,
               pathway_new = "WGD",
               Tumor_Sample_Barcode = sample_id,
               aliquot_id = sample_id,
               event = "WGD",
               Hugo_Symbol = "WGD") %>%
  filter(!is.na(wgd_mediantiming)) 
  
setDT(wgd_median)
wgd_median[, (paste0("X", 0:249)) := lapply(1:250, \(x) wgd_mediantiming)]

wgd_median <- wgd_median %>% mutate(Hugo_Symbol = "WGD",
         Chromosome = NA, 
         Start_Position = NA, 
         End_Position= NA, 
         Reference_Allele = NA, 
         Tumor_Seq_Allele1 = NA,
         Tumor_Seq_Allele2 = NA,
         Variant_Classification = "Missense_Mutation", 
         Variant_Type = NA) %>%  
  mutate(median_value = rowMedians(as.matrix(dplyr::select(., all_of(column_names))), na.rm = TRUE),
         early_ratio = rowMeans(dplyr::select(., all_of(column_names)) < 0.5, na.rm = TRUE)) %>%
  dplyr::select(all_of(kept_cols))
write.csv(wgd_median, file = paste0(path_gritic_pathway, "/diff_WGD.csv"), row.names = F)
```

# 10. Median timing
## Median for all indexes
```{r}
data_timing_median <- data_timing_ratio %>%
  filter(Tumor_Sample_Barcode %in% sample_sucess) %>%
  filter(Tumor_Sample_Barcode %in% sample_MSS) %>%
  mutate(aliquot_id = Tumor_Sample_Barcode) %>%
  group_by(aliquot_id) %>%
  slice_min(
    abs(median_value - median(median_value, na.rm = TRUE)),
    n = 1,
    with_ties = FALSE
  ) %>%
  ungroup() %>%
  mutate(pathway = "Median",
         event = "Median",
         Hugo_Symbol = "Median",
         Other_Info = NA,
         pathway_new = "Median",
         Chromosome = NA, 
         Start_Position = NA, 
         End_Position= NA, 
         Reference_Allele = NA, 
         Tumor_Seq_Allele1 = NA,
         Tumor_Seq_Allele2 = NA,
         Variant_Classification = "Missense_Mutation", 
         Variant_Type = NA,
         Tumor_Sample_Barcode = aliquot_id,
         sample_id = aliquot_id) %>%
  dplyr::select(all_of(kept_cols)) 
write.csv(data_timing_median, file = paste0(path_gritic_pathway, "/diff_Median_index.csv"), row.names = F)
```

# 11. Combine all timing results
## Pathways without HLALOH
```{r}
## Driver mutations
data_timing_driver <- fread(paste0(path_gritic_pathway, "/diff_drivermutations.csv")) %>% 
  filter(Tumor_Sample_Barcode %in% sample_sucess) %>%
  filter(Tumor_Sample_Barcode %in% sample_MSS) %>%
  mutate(driver_gene = "yes") %>% 
  mutate(aliquot_id = Tumor_Sample_Barcode)

## APM
data_timing_APM_all <- fread(paste0(path_gritic_pathway, "/diff_APM.csv")) %>% 
                       mutate(regulator = "other",
                              aliquot_id = Tumor_Sample_Barcode,
                              driver_gene = "no")
data_timing_pathway_all <- fread(paste0(path_gritic_pathway, "/diff_all.csv")) %>% 
                       mutate(aliquot_id = Tumor_Sample_Barcode,
                              driver_gene = "no")

## Median
data_timing_median <- fread(paste0(path_gritic_pathway, "/diff_Median_index.csv")) %>% 
                       mutate(aliquot_id = Tumor_Sample_Barcode)

diff_all <- rbind(
  as.data.table(data_timing_driver),
  as.data.table(data_timing_APM_all),
  as.data.table(data_timing_pathway_all),
  as.data.table(data_timing_median),
  fill = TRUE) %>%   
  filter(Tumor_Sample_Barcode %in% sample_sucess) %>%
  filter(Tumor_Sample_Barcode %in% sample_MSS)

diff_all_selected  <- diff_all %>%
  group_by(pathway, histology_abbreviation) %>%
  summarise(n_samples = n_distinct(aliquot_id), .groups = "drop") %>%
  ungroup() %>% 
  inner_join(diff_all, ., by = c("pathway", "histology_abbreviation")) %>%
  ungroup()
write.csv(diff_all_selected, paste0(path_gritic_pathway, "diff_all_selected_drivermutation.csv"), row.names = F)

check <- diff_all %>% filter(pathway == "Median")
check <- diff_all_selected %>% filter(pathway == "Median")
check <- diff_all %>% filter(pathway == "Median")
```

## Pathways with HLALOH
```{r}
sample_wgd = unique(wgd_median$sample_id)

## Same
data_timing_wgd_all <- fread(paste0(path_gritic_pathway, "/diff_WGD.csv")) %>% 
                       mutate(aliquot_id = Tumor_Sample_Barcode,
                              driver_gene = "no",
                              sample_id = Tumor_Sample_Barcode)

data_timing_loss <- fread(paste0(path_gritic_pathway, "/diff_loss.csv")) %>% 
                       mutate(aliquot_id = Tumor_Sample_Barcode,
                              driver_gene = "no")

diff_all <- rbind(
  as.data.table(data_timing_driver),
  as.data.table(data_timing_APM_all),
  as.data.table(data_timing_pathway_all),
  as.data.table(data_timing_median),
  as.data.table(data_timing_wgd_all),
  as.data.table(data_timing_loss), fill = TRUE) %>% 
  mutate(aliquot_id = sample_id) %>%   
  filter(Tumor_Sample_Barcode %in% sample_sucess) %>%
  filter(Tumor_Sample_Barcode %in% sample_MSS) %>%
  filter(Tumor_Sample_Barcode %in% sample_wgd)

diff_all_selected <- diff_all %>%
  filter(Tumor_Sample_Barcode %in% sample_wgd) %>%
  group_by(pathway, histology_abbreviation) %>%
  summarise(n_samples = n_distinct(aliquot_id), .groups = "drop") %>%
  ungroup() %>% 
  inner_join(diff_all, ., by = c("pathway", "histology_abbreviation")) %>%
  ungroup()
write.csv(diff_all_selected, paste0(path_gritic_pathway, "diff_all_selected_loss.csv"), row.names = F)

## Check results
check <- diff_all_selected %>% filter(histology_abbreviation == "Liver-HCC", pathway == "HLALOH")

check <- data_timing_pathway_all %>% 
  filter(pathway == "Cellular Amino Acid Metabolic Process") %>%
  filter(histology_abbreviation %in% c("Breast-AdenoCA", "ColoRect-AdenoCA"))

check <- diff_all_selected  %>% 
  filter(pathway == "Cellular Amino Acid Metabolic Process") %>%
  filter(histology_abbreviation %in% c("Breast-AdenoCA", "ColoRect-AdenoCA"))

check <- diff_all_selected  %>% 
  filter(pathway == "HLALOH") %>%
  filter(histology_abbreviation %in% c("Breast-AdenoCA", "ColoRect-AdenoCA"))
```

# 12. Bootstrap
## 1000 bootstraps
### Pathway min and driver mutation
```{r}
# Driver mutation
path_save = "/Users/wchen20/Desktop/PCAWG/MHC_evolution/GRITIC/output/allmutation/mafgritic_nonsyn/pathway/min_drivermutation_250/"
path_plot= "/Users/wchen20/Desktop/2026-02-16_Draft/Revision/R1Major5/pathway_min/min_5/min_drivermutation/"
```

## 1000 bootstraps without median
```{r}
replacement = "yes"
median_reference = "no"
n_boot = 1000
savename = "min"
var = "median_value"
direction = "lower"
number = 5
number_sk = 5
use_pro = "yes"
```

## Read data
```{r}
included_cnv = "wgd"
included_cnv = "wgd_hlaloh"
included_cnv = "nocnv"

if (included_cnv == "wgd_hlaloh") {
  diff_all_selected <- fread(paste0(path_save, "diff_all_selected_loss.csv")) %>% 
  filter(sample_id %in% sample_wgd)
  
  wgd_type = inner_join(wgd_median, LOH_type, by = "sample_id")
  freq_type <- as.data.frame(table(wgd_type$histology_abbreviation))
  names(freq_type) <- c("histology_abbreviation", "type_count")

} else if (included_cnv == "wgd") {
  diff_all_selected <- fread(paste0(path_save, "diff_all_selected_HLALOH.csv")) %>% 
    filter(pathway != "HLALOH")
} else if (included_cnv == "nocnv") {
   diff_all_selected <- fread(paste0(path_save, "diff_all_selected_drivermutation.csv"))
   
   sample_unique <- diff_all_selected %>% distinct(sample_id, .keep_all = T)
   freq_type <- as.data.frame(table(sample_unique$histology_abbreviation))
   names(freq_type) <- c("histology_abbreviation", "type_count")
}
```

# 13. Timeline
```{r}
n_pathway = 3
n_driver_included = 5
n_pathway_included = 10
types = unique(diff_all_selected$histology_abbreviation)
only_table = "yes"

for (type in types) {
  
  if (!type %in% c("Skin-Melanoma")) {
    number_filter = number
  } else {
    number_filter = number_sk
  }
  
  cat(type, "\n")
  
  drivergene_cancer <- driver_list %>% filter(grepl(type, Tissue)) %>% pull(Gene)
  
  diff_all_selected_driver <- diff_all_selected %>% 
    filter(driver_gene == "yes") %>%
    filter(pathway %in% drivergene_cancer)
  
  diff_all_selected_pathway <- diff_all_selected %>% 
    filter(driver_gene == "no")
  
  diff_all_selected_median <- diff_all_selected %>% 
    filter(pathway == "Median")
  
  diff_all_included_sub <- rbind(diff_all_selected_driver, diff_all_selected_pathway) %>%
    rbind(., diff_all_selected_median) %>%
    filter(histology_abbreviation == !!type) %>% 
    filter(n_samples >= !!number_filter) %>%
    group_by(sample_id) %>%
    filter(n_distinct(pathway) >= n_pathway) %>%
    ungroup()
    
  top_genes <- diff_all_included_sub %>%
  filter(driver_gene == "yes") %>%
  inner_join(freq_type, by = "histology_abbreviation") %>%
  mutate(prevelance = n_samples / type_count) %>%
  filter(prevelance > 0.05) %>%
  distinct(pathway, .keep_all = T) %>%
  group_by(histology_abbreviation) %>%
  dplyr::slice_max(
    order_by = n_samples,
    n = n_driver_included,
    with_ties = F
  ) %>%
  ungroup() %>%
  pull(pathway) %>%
  unique()
  
  top_pathways <- diff_all_included_sub %>%
  filter(driver_gene != "yes") %>%
  filter(pathway != "yes") %>%
  inner_join(freq_type, by = "histology_abbreviation") %>%
  mutate(prevelance = n_samples / type_count) %>%
  filter(prevelance > 0.05) %>%
  distinct(pathway, .keep_all = T) %>%
  group_by(histology_abbreviation, pathway) %>%
  dplyr::slice_max(
    order_by = n_samples,
    n = n_pathway_included,
    with_ties = T) %>%
  ungroup() %>%
  pull(pathway) %>%
  unique()
  
  if (median_reference == "yes") {
    
    diff_all_included_sub <- diff_all_selected %>% 
        filter(histology_abbreviation == !!type) %>% 
        filter(n_samples >= !!number_filter) %>%
        filter(pathway %in% c(top_genes, top_pathways, "Median")) %>%
        group_by(sample_id) %>%
        filter(n_distinct(pathway) >= n_pathway) %>%
        ungroup() 
     
     ref_cat = "Median"
   
  } else if (median_reference == "no") {
    
      diff_all_included_sub <- diff_all_selected %>% 
        filter(histology_abbreviation == !!type) %>% 
        filter(n_samples >= !!number_filter) %>%
        filter(pathway %in% c(top_genes, top_pathways, "Median")) %>%
        group_by(sample_id) %>%
        filter(n_distinct(pathway) >= n_pathway) %>%
        ungroup() %>%
        filter(pathway != "Median")
      
      ref_cat <- diff_all_included_sub %>% arrange(pathway) %>%
        dplyr::slice(1) %>%
        pull(pathway)
      
  }
  
  if (nrow(diff_all_included_sub) > 0) {
      setDT(diff_all_included_sub)
      table_prob <- generate_pairwise(dt = diff_all_included_sub,
                                      sample_col = "sample_id",
                                      event_col = "pathway")
      
      dir.create(paste0(path_plot, "/PhylogicNDT/", included_cnv, "/", type, "/"), recursive = T)
      
      fwrite(
        table_prob,
        file = file.path(paste0(path_plot, "/PhylogicNDT/", included_cnv, "/", type, "/", type, ".comp.tsv")),
        sep = "\t")
         
      samplings <- c("all", 50, 100, 150, 200)
      
      for (sampling in samplings) {
  
      path_comp_all = paste0(path_plot, "/PhylogicNDT/", included_cnv, "/", type, "_", sampling, "/")
      dir.create(path_comp_all, recursive = T)
      
      table_prob <- compare_event_timing(dt = diff_all_included_sub,
                                         sample_col = "aliquot_id",
                                         event_col = "pathway",
                                         n_cols = sampling,
                                         seed = 123)

      fwrite(
        table_prob,
        file = file.path(path_comp_all, paste0(type, ".comp.tsv")),
        sep = "\t")
      }
      
  if (only_table == "yes") {
    cat(paste0("Outputing for ",type, "\n"))
  } else {
  
  bootstrap_results <- vector("list", n_boot)
  set.seed(123)
  sample_ids_all <- unique(diff_all_included_sub$aliquot_id)

  bootstrap_results <- list()

  if (replacement %in% c("yes", "no")) {
    
    for (b in seq_len(n_boot)) {

    if (replacement == "no") {

      sampled_ids <- sample(
        sample_ids_all,
        size = ceiling(length(sample_ids_all) * 0.8),
        replace = FALSE
      )

      boot_data <- diff_all_included_sub %>%
        filter(aliquot_id %in% sampled_ids)

    } else {

      sampled_ids <- sample(
        sample_ids_all,
        replace = TRUE
      )

      boot_cols <- paste0("X", 0:249)

      boot_data <- lapply(seq_along(sampled_ids), function(i) {

        sid <- sampled_ids[i]

        tmp <- diff_all_included_sub %>%
          filter(aliquot_id == sid)

        tmp$aliquot_id <- paste0(tmp$aliquot_id, "_boot", i)

        chosen_col <- sample(boot_cols, 1)

        tmp$median_value <- tmp[[chosen_col]]

        tmp <- tmp %>%
          group_by(pathway, aliquot_id) %>%
          dplyr::slice_min(median_value, n = 1) %>%
          ungroup()

        tmp
      }) %>%
        bind_rows()
    }
    
    path_comp = paste0(path_plot, "/comp/", type, "/", b, "/")
    dir.create(path_comp, recursive = T)
    
    if (use_pro == "no") {
          
      temp_final <- generate_winner_table(
      data          = boot_data,
      gene_col      = "pathway",
      sample_col    = "aliquot_id",
      histology_col = "histology_abbreviation",
      value_col     = var,
      order         = direction,
      path_comp     = path_comp,
      comp_median   = "yes") %>% 
      filter(!(Gene1.wins == 0 & Gene2.wins == 0))
      
    } else if (use_pro == "yes") {
      setDT(boot_data)
      table_prob <- generate_pairwise(dt = boot_data,
                                      sample_col = "aliquot_id",
                                      event_col = "pathway")
      
      fwrite(table_prob,
             file = file.path(paste0(path_comp, b, ".comp.tsv")),
             sep = "\t")
      
      temp_final <- table_prob %>%
        mutate(Gene1.win = as.integer(`p1->2` > `p2->1`),
               Gene2.win = as.integer(`p2->1` > `p1->2`),
               Unknown   = as.integer(`p1->2` == `p2->1`)) %>%
        group_by(eve1, eve2) %>%
        summarise(Gene1.wins = sum(Gene1.win),
                  Gene2.wins = sum(Gene2.win),
                  unknown    = sum(Unknown),
                  .groups = "drop") %>%
        rename(Mutation1.Gene = eve1,
               Mutation2.Gene = eve2) %>% 
        filter(!(Gene1.wins == 0 & Gene2.wins == 0)) %>% 
        filter(`Mutation1.Gene` != `Mutation2.Gene`)
      
    }

    if (nrow(temp_final) == 0) next

    all_players <- sort(unique(c(
      temp_final$Mutation1.Gene,
      temp_final$Mutation2.Gene
    )))

    temp_final <- temp_final %>%
      mutate(
        Mutation1.Gene = factor(
          Mutation1.Gene,
          levels = all_players
        ),
        Mutation2.Gene = factor(
          Mutation2.Gene,
          levels = all_players
        )
      )

    BT_results <- tryCatch({
          BTm(
          cbind(Gene1.wins, Gene2.wins),
          Mutation1.Gene,
          Mutation2.Gene,
          refcat = ref_cat,
          data = temp_final
        )

    }, error = function(e) NULL)

    if (is.null(BT_results)) next

    bt_df <- as.data.frame(summary(BT_results)$coefficients)

    bt_df$Gene <- substring(rownames(bt_df), 3)
    bt_df$bootstrap <- b

    bootstrap_results[[b]] <- bt_df
  }

  } else {

  for (b in seq_along(column_names)) {

    col <- column_names[b]
    boot_data <- copy(diff_all_included_sub)
    boot_data$median_value <- boot_data[[col]]
    boot_data <- boot_data %>%
          group_by(pathway, sample_id) %>%
          slice_min(median_value, n = 1) %>%
          ungroup()
    
    path_comp = paste0(path_plot, "/comp/", type, "/", b, "/")
    dir.create(path_comp, recursive = T)
    
    temp_final <- generate_winner_table(
      data          = boot_data,
      gene_col      = "pathway",
      sample_col    = "aliquot_id",
      histology_col = "histology_abbreviation",
      value_col     = var,
      order         = direction,
      path_comp     = path_comp,
      comp_median   = "yes"
    ) %>% filter(!(Gene1.wins == 0 & Gene2.wins == 0))

    if (nrow(temp_final) == 0) next

    all_players <- sort(unique(c(
      temp_final$Mutation1.Gene,
      temp_final$Mutation2.Gene
    )))

    temp_final <- temp_final %>%
      mutate(
        Mutation1.Gene = factor(
          Mutation1.Gene,
          levels = all_players
        ),
        Mutation2.Gene = factor(
          Mutation2.Gene,
          levels = all_players
        )
      )

    BT_results <- tryCatch({

         BTm(
          cbind(Gene1.wins, Gene2.wins),
          Mutation1.Gene,
          Mutation2.Gene,
          refcat = ref_cat,
          data = temp_final
        )

    }, error = function(e) NULL)

    if (is.null(BT_results)) next

    bt_df <- as.data.frame(summary(BT_results)$coefficients)
    bt_df$Gene <- substring(rownames(bt_df), 3)
    bt_df$bootstrap <- b
    bootstrap_results[[col]] <- bt_df
  }
  }

  bootstrap_results_combined <- bind_rows(bootstrap_results) %>% mutate(bootstrap = as.integer(bootstrap))
  ref_row <- tibble(Estimate = rep(0, n_boot),
                      `Std. Error` = rep(NA_real_),
                      `z value` = rep(NA_real_),
                      `Pr(>|z|)` = rep(NA_real_),
                      Gene = ref_cat,
                      bootstrap = 1:n_boot)
      
  bootstrap_results_df <- bind_rows(ref_row, bootstrap_results_combined)
  
  if (nrow(bootstrap_results_df) > 100) {
    
  bootstrap_results_df$Gene <- sapply(bootstrap_results_df$Gene, insert_newlines_every_five_words)
  
  width = 5.5
  height = 6
  plot_BT(bootstrap_results_df, path_plot, type, width, height)
  }
  }
  }
}
```