---
title: "R1Major5_uncertainty_wgd_tp53"
author: "AJ"
date: "2026-04-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Library packages
```{r}
rm(list = ls())

suppressWarnings(library("stringr"))
suppressWarnings(library("dplyr"))
suppressWarnings(library("patchwork"))
suppressWarnings(library("ggpubr"))
suppressWarnings(library("ggridges"))
suppressWarnings(library("truncnorm"))
suppressWarnings(library("tidyr"))
suppressWarnings(library("scatterpie"))
suppressWarnings(library("tidyselect"))
suppressWarnings(library("pillar"))
suppressWarnings(library("data.table"))
suppressWarnings(library("readr"))
suppressWarnings(library("ggplot2"))
suppressWarnings(library("openxlsx"))
suppressWarnings(library("ggrepel"))
suppressWarnings(library("tibble"))
suppressWarnings(library("textshape"))
suppressWarnings(library("matrixStats"))
suppressWarnings(library("bayestestR"))

## Path
PCAWG_path <- "/rsrch6/scratch/hema_bio-Malignan/wchen20/PCAWG"
PCAWG_path <- "/Users/wchen20/Desktop/PCAWG"

source("/Users/wchen20/Desktop/PCAWG/code/source/basic.R")
source("/Users/wchen20/Desktop/PCAWG/code/source/insert_newlines.R")
source("/Users/wchen20/Desktop/PCAWG/code/source/plot_segTiming.R")

## Timing of WGD
wgd_median <- fread("/Users/wchen20/Desktop/PCAWG/MHC_evolution/GRITIC/output/wgd_median.csv") %>% dplyr::select(wgd_mediantiming, sample_id)

path_hdi = "/Users/wchen20/Desktop/PCAWG/MHC_evolution/GRITIC/output/allmutation/all_HDI/"
path_save = "/Users/wchen20/Desktop/2026-02-16_Draft/Revision/R1Major5/"

cols <- paste0("X", 0:249)
breaks_seq <- seq(0, 0.95, by = 0.05)
bin_levels <- c(
  levels(cut(0, breaks = breaks_seq, include.lowest = TRUE, right = FALSE)),
  "> 0.95"
)

breaks_seq <- seq(0, 1.05, by = 0.05)
breaks_seq_nrpcc <- seq(0, 80, by = 5)

plot_uncertainty <- function(data) {
  
  df_freq <- data %>%
    dplyr::count(uncertainty_bin) %>%
    mutate(proportion = n / sum(n))
  
  plot <- ggplot(df_freq, aes(x = uncertainty_bin, y = proportion)) +
  geom_bar(stat = "identity", fill = "#00A087") +
  labs(
    x = "",
    y = "Proportion",
    title = "Distribution of Uncertainty"
  ) +
  theme_bw() +
  theme(axis.title.y = element_text(size = 10, colour = "black"),
        axis.title.x = element_text(size = 14),
        legend.title = element_blank(), 
        legend.text = element_text(size = 12),
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 8, colour = "black"),
        axis.text.y = element_text(size = 8, colour = "black"),
        title = element_text(size = 10),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank()) +
  coord_cartesian(ylim = c(0, 0.6))
  
  return(plot)
  
}
```

# Set path
```{r}
path_figure <- "/Users/wchen20/Desktop/2026-02-16_Draft/Revision_Figures/Original_output/"

path_gritic = "/Users/wchen20/Desktop/PCAWG/MHC_evolution/GRITIC/output/allmutation/mafgritic_nonsyn/"
path_gritic_pathway = "/Users/wchen20/Desktop/2026-02-16_Draft/Final_Results/8_MutationTiming/min_drivermutation_250/"

path_HLALOH = "/Users/wchen20/Desktop/2026-02-16_Draft/Final_Results/1_APM/2_HLALOH/"
```

# 1. Check HDI
## Manul
```{r}
data_timing_driver_all <- fread(paste0(path_gritic_pathway, "/nonsyn_filter_diff.csv"))

check_250 <- data_timing_driver_all %>% 
             filter(Hugo_Symbol == "HIVEP3",
                    Chromosome == 1,
                    Start_Position == 41978926,
                    End_Position == 41978926,
                    Variant_Classification == "Missense_Mutation",
                    Tumor_Sample_Barcode == "5975deb5-00d2-4d18-b5c0-34ac9947fe11")

## Check HDI
p = 0.95
hdi_results <- t(apply(
  as.matrix(check_250[, ..column_names]),
  1,
  function(x) {
    h <- hdi(x, ci = p)
    c(
      HDI_low  = h$CI_low,
      HDI_high = h$CI_high
    )
  }
))

check_250[, `:=`(
  HDI_low  = hdi_results[, "HDI_low"],
  HDI_high = hdi_results[, "HDI_high"],
  hdi_width_check = hdi_results[, "HDI_high"] - hdi_results[, "HDI_low"]
)]

check_250
```

## 2. 95% HDI
```{r}
i = 5
mafgritic_files <- list.files(path = path_hdi, pattern = paste0("*_", i, ".csv"), recursive = FALSE, full.names = TRUE)
  
data_all <- lapply(mafgritic_files, fread) %>% 
    rbindlist() %>% 
    mutate(sample_id = Tumor_Sample_Barcode) %>% 
    inner_join(LOH_type, by = "sample_id") 

Check_HDI <- data_all %>% 
             filter(Hugo_Symbol == "HIVEP3",
                    Chromosome == 1,
                    Start_Position == 41978926,
                    End_Position == 41978926,
                    Variant_Classification == "Missense_Mutation",
                    Tumor_Sample_Barcode == "5975deb5-00d2-4d18-b5c0-34ac9947fe11")
```

# Uncertainty by WGD and TP53
```{r}
props <- c(0.5, 0.6, 0.7, 0.8, 0.95)
plot_uncertainty_all_list <- list()
plot_uncertainty_wgd_list <- list()
plot_uncertainty_nowgd_list <- list()
plot_uncertainty_tp53_list <- list()
plot_uncertainty_tp53_ov_list <- list()

for (i in props) {
  
  p <- props[i]

  mafgritic_files <- list.files(path = path_hdi, pattern = paste0("*_", i, ".csv"), recursive = FALSE, full.names = TRUE)
  
  data_all <- lapply(mafgritic_files, fread) %>% 
    rbindlist() %>% 
    mutate(sample_id = Tumor_Sample_Barcode) %>% 
    inner_join(LOH_type, by = "sample_id") 
  plot_uncertainty_all_list[[i]] <- plot_uncertainty(data_all)
  
  data_all_clonal <- data_all %>% filter(subclonal_ratio < 0.5) 
  
  data_nonsyn <- lapply(mafgritic_files, fread) %>% 
    rbindlist() %>% 
    mutate(sample_id = Tumor_Sample_Barcode) %>% 
    inner_join(LOH_type, by = "sample_id") %>%
    filter(Variant_Classification %in% non_syn)
  
  plot_uncertainty_nonsyn_list[[i]] <- plot_uncertainty(data_nonsyn)
  
  data_wgd <- data_all %>% 
    mutate(sample_id = Tumor_Sample_Barcode) %>% 
    inner_join(., wgd_median, by = "sample_id")
  
  plot_uncertainty_wgd_list[[i]] <- plot_uncertainty(data_wgd)
  
  data_nowgd <- data_all %>% filter(!Tumor_Sample_Barcode %in% unique(wgd_median$sample_id))
  plot_uncertainty_nowgd_list[[i]] <- plot_uncertainty(data_nowgd)
  
  data_tp53 <- data_all %>%
  filter(Hugo_Symbol == "TP53", Variant_Classification %in% non_syn)
  plot_uncertainty_tp53_list[[i]] <- plot_uncertainty(data_tp53)
  
  data_tp53_ov <- data_all %>%
  filter(Hugo_Symbol == "TP53", Variant_Classification %in% non_syn, histology_abbreviation == "Ovary-AdenoCA")
  plot_uncertainty_tp53_ov_list[[i]] <- plot_uncertainty(data_tp53_ov)

  }

# Uncertainty for different intervals
plot_uncertainty_all_com <- (plot_uncertainty_all_list[[1]] + plot_uncertainty_all_list[[2]])/(plot_uncertainty_all_list[[3]] + plot_uncertainty_all_list[[4]])
  
ggsave(
  filename = paste0(path_save, "/HDI_uncertainty_all_CIs.pdf"),
  plot     = plot_uncertainty_all_com,
  width = 12, 
  height = 5.5,
  units    = "in")

# Uncertainty for 95%CIs intervals
i = 5 

## All mutations  
ggsave(
  filename = paste0(path_save, "/HDI_uncertainty_95_CIs.pdf"),
  plot     = plot_uncertainty_all_list[[i]],
  width = 6, 
  height = 2.5,
  units    = "in")

## All Nonsynomynous mutations 
plot_uncertainty_nonsyn_list
ggsave(filename = paste0(path_save, "/HDI_uncertainty_nonsyn_95_CIs.pdf"),
       plot     = plot_uncertainty_nonsyn_list[[i]],
       width = 6, 
       height = 2.5,
       units    = "in")

# Uncertainty for different intervals by WGD
plot_uncertainty_wgd_com <- (plot_uncertainty_wgd_list[[1]] + plot_uncertainty_wgd_list[[2]])/(plot_uncertainty_wgd_list[[3]] + plot_uncertainty_wgd_list[[4]])
  
ggsave(
  filename = paste0(path_save, "/HDI_uncertainty_wgd_CIs.pdf"),
  plot     = plot_uncertainty_wgd_com,
  width = 12, 
  height = 5.5,
  units    = "in")
  
plot_uncertainty_nonwgd_com <- (plot_uncertainty_nowgd_list[[1]] + plot_uncertainty_nowgd_list[[2]])/(plot_uncertainty_nowgd_list[[3]] + plot_uncertainty_nowgd_list[[4]])
  
ggsave(
  filename = paste0(path_save, "/HDI_uncertainty_nowgd_CIs.pdf"),
  plot     = plot_uncertainty_nonwgd_com,
  width = 12, 
  height = 5.5,
  units    = "in")

plot_uncertainty_all_com <- plot_uncertainty_all_list[[5]] + plot_uncertainty_wgd_list[[5]] + plot_uncertainty_nowgd_list[[5]]
ggsave(filename = paste0(path_save, "/HDI_uncertainty_wgd_CIs.pdf"),
       plot     = plot_uncertainty_all_com,
       width = 20, 
       height = 5,
       units    = "in")

# Uncertainty for different intervals by TP53
plot_uncertainty_tp53_com <- (plot_uncertainty_tp53_list[[1]] + plot_uncertainty_tp53_list[[2]])/(plot_uncertainty_tp53_list[[3]] + plot_uncertainty_tp53_list[[4]])
  
ggsave(
  filename = paste0(path_save, "/HDI_uncertainty_tp53_CIs.pdf"),
  plot     = plot_uncertainty_tp53_com,
  width = 12, 
  height = 5.5,
  units    = "in")
  
plot_uncertainty_all_com <- plot_uncertainty_all_list[[5]] + plot_uncertainty_tp53_list[[5]] + plot_uncertainty_tp53_ov_list[[5]] 
ggsave(filename = paste0(path_save, "/HDI_uncertainty_tp53_CIs.pdf"),
       plot     = plot_uncertainty_all_com,
       width = 20, 
       height = 5,
       units    = "in")

plot_uncertainty_all_com <- plot_uncertainty_tp53_list[[i]] + plot_uncertainty_tp53_ov_list[[i]]
ggsave(filename = paste0(path_save, "/HDI_uncertainty_tp53_CIs.pdf"),
       plot     = plot_uncertainty_all_com,
       width = 12, 
       height = 3,
       units    = "in")
```

# Uncertainty by CNV
```{r}
path_hdi <- "/Users/wchen20/Desktop/PCAWG/MHC_evolution/GRITIC/output/allmutation/all_HDI/"
i = 5
mafgritic_files <- list.files(path = path_hdi, pattern = paste0("*_", i, ".csv"), recursive = FALSE, full.names = TRUE)

df_snv  <- lapply(mafgritic_files, fread) %>% 
  rbindlist() %>% 
  mutate(sample_id = Tumor_Sample_Barcode,
         Chromosome = as.character(Chromosome),
         Start_Position = as.numeric(Start_Position),
         End_Position = as.numeric(End_Position)) %>% 
  inner_join(LOH_type, by = "sample_id") %>% 
  filter(subclonal_ratio < 0.5) 
#write.csv(df_snv, file = paste0(path_hdi, "hdi_results_bin_5_all.csv"), row.names = F)

#df_snv <- fread(paste0(path_hdi, "hdi_results_bin_5_all.csv"))

## CNV
cnv_hd <- fread("/Users/wchen20/Desktop/PCAWG/consensus_cnv/somatic_cna_annotated_merged.txt") %>%
  mutate(total_cn = as.numeric(total_cn),
         major_cn = as.numeric(major_cn),
         minor_cn = as.numeric(minor_cn),
         start = as.numeric(start),
         end = as.numeric(end)) %>%
  filter(!is.na(major_cn), !is.na(minor_cn)) %>%
  filter((major_cn + minor_cn) < 10) %>%
  mutate(chromosome = paste0("chr", gsub("^chr", "", chromosome))) %>%
  filter(chromosome %in% paste0("chr", 1:22)) %>% 
  dplyr::select("chromosome", "start", "end", "total_cn", "major_cn", "minor_cn", "sample_id") %>% 
  mutate(chromosome = sub("^chr", "", chromosome),
         cnv_group = paste0(major_cn, " + ", minor_cn)) %>% 
  filter(!(major_cn==0 & minor_cn==0)) %>%
  filter(!(major_cn==1 & minor_cn==2)) %>%
  filter(!(major_cn==0 & minor_cn==1))

## NRPCC
NRPCC <- fread("/Users/wchen20/Desktop/PCAWG/evolution_and_heterogeneity/icgc_sample_annotations_summary_table.txt") %>% 
  mutate(sample_id = tumour_aliquot_id) %>% 
  mutate(nrpcc_group = ifelse(nrpcc > 10, "NRPCC > 10", "NRPCC =< 10"),
         nrpcc_bin = cut(nrpcc, breaks = breaks_seq_nrpcc, include.lowest = TRUE, right = FALSE),
         purity_bin = cut(purity, breaks = breaks_seq, include.lowest = TRUE, right = FALSE)) %>% 
  filter(sample_id %in% sample_MSS)

## Combined
df_merged_NRPCC <- df_snv %>%
  left_join(cnv_hd, 
            by = join_by(
              Tumor_Sample_Barcode == sample_id,
              Chromosome == chromosome,
              Start_Position >= start,
              Start_Position <= end
            )) %>%
  mutate(total_cn = as.character(total_cn),
         total_cn = case_when(major_cn == 1 & minor_cn == 1 ~ "noCNV",
                              is.na(total_cn) ~ "noCNV",
                              TRUE ~ total_cn),
         total_cn = factor(total_cn, levels = c("noCNV", "0", "1", "2", "3", "4", "5", "6", "7", "8","9"))) %>% 
  left_join(., wgd_median, by = "sample_id") %>%
  mutate(wgd = ifelse(!is.na(wgd_mediantiming), "WGD", "NonWGD")) %>%
  filter(!is.na(cnv_group)) %>% 
  inner_join(., NRPCC, by = "sample_id")
```

## By Total copy number
```{r}
## All uncertainty         
plot_un <- ggplot(df_merged_NRPCC, aes(x = factor(total_cn), y = hdi_width)) +
  geom_violin(trim = FALSE, fill = "#00A087", alpha = 0.6) +
  geom_boxplot(width = 0.1, outlier.shape = NA, fill = "white", alpha = 0.8) +
  theme_bw() +
  labs(x = "Total Copy Number", y = "Uncertainty", title = "Distribution of Uncertainty (CI = 0.95)") +
  theme(
    axis.text.x = element_text(vjust = 1, size = 14, colour = "black"),
    axis.text.y = element_text(size = 14, colour = "black"),
    title = element_text(size = 16),
    axis.title  = element_text(size = 15),
    plot.margin = margin(10, 10, 10, 10),
  )

ggsave(filename = paste0(path_save, "/HDI_uncertainty_cnv.pdf"),
       plot     = plot_un,
       width = 12, 
       height = 4,
       units    = "in")

### by WGD
plot_un <- ggplot(df_merged_NRPCC, aes(x = factor(total_cn), y = hdi_width)) +
  geom_violin(trim = FALSE, fill = "#00A087", alpha = 0.6) +
  geom_boxplot(width = 0.1, outlier.shape = NA, fill = "white", alpha = 0.8) +
  facet_wrap(~ wgd, ncol = 1, ) +
  theme_bw() +
  labs(x = "Total Copy Number", y = "Uncertainty", title = "Distribution of Uncertainty (CI = 0.95)") +
  theme(
    axis.text.x = element_text(vjust = 1, size = 14, colour = "black"),
    axis.text.y = element_text(size = 14, colour = "black"),
    axis.title  = element_text(size = 16),
    title = element_text(size = 16),
    plot.margin = margin(10, 10, 10, 10),
    strip.text  = element_text(size = 15),
  )

ggsave(filename = paste0(path_save, "/HDI_uncertainty_wgd.pdf"),
       plot     = plot_un,
       width = 12, 
       height = 7.5,
       units    = "in")
```

# By copy number combinations
## All CNV
```{r}
plot_un <- ggplot(df_merged_NRPCC, aes(x = factor(cnv_group), y = hdi_width)) +
  geom_violin(trim = FALSE, width = 2, fill = "#00A087", alpha = 0.6) +
  geom_boxplot(width = 0.1, outlier.shape = NA, fill = "white", alpha = 0.8) +
  facet_wrap(~ wgd, ncol = 1, ) +
  theme_bw() +
  labs(x = "", y = "Uncertainty", title = "Distribution of Uncertainty (CI = 0.95)") +
  theme(
    axis.text.x = element_text(vjust = 1, hjust = 1, size = 14, colour = "black", angle = 45),
    axis.text.y = element_text(size = 14, colour = "black"),
    axis.title  = element_text(size = 16),
    title = element_text(size = 16),
    plot.margin = margin(10, 10, 10, 10),
    strip.text  = element_text(size = 15),
  ) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.2), limits = c(0, 1))

ggsave(filename = paste0(path_save, "/HDI_uncertainty_wgd_cnvcom.pdf"),
       plot     = plot_un,
       width = 12, 
       height = 7.5,
       units    = "in")

##
data_plot <- df_merged_NRPCC %>% filter(!cnv_group %in% c("1 + 0", "1 + 1"))

plot_un <- ggplot(
  data_plot,
  aes(x = factor(wgd), y = hdi_width, fill = factor(wgd))) +
  geom_violin(trim = FALSE, alpha = 0.6) +
  geom_boxplot(
    width = 0.15,
    outlier.shape = NA,
    fill = "white"
  ) +
  facet_wrap(~cnv_group, nrow = 2) +
  stat_compare_means(
    method = "wilcox.test",
    label = "p.signif",
    label.y = 1.05,
    label.x = 1.3,
    size = 4
  ) +
  theme_bw() +
  labs(
    x = "",
    y = "Uncertainty"
  ) + 
  theme(
    axis.text.x = element_blank(),
    axis.text.y = element_text(size = 10, colour = "black"),
    axis.title  = element_text(size = 12),
    title = element_text(size = 12),
    plot.margin = margin(10, 10, 10, 10),
    strip.text  = element_text(size = 12),
    legend.position = "top"
  ) + 
  scale_y_continuous(breaks = seq(0, 1, by = 0.2), limits = c(0, 1.1)) +  
  scale_fill_manual(values = c("#F39B7F", "#8491B4"),
                    name = "WGD")

ggsave(filename = paste0(path_save, "/HDI_uncertainty_wgd_cnvcom_wilcox.pdf"),
       plot     = plot_un,
       width = 12, 
       height = 6,
       units    = "in")
```

### Samples with WGD timing in 0.4~0.6
```{r}
df_merged_sub <- df_merged_NRPCC %>% filter(wgd_mediantiming >= 0.4, wgd_mediantiming <= 0.6)
check <- df_merged_sub %>% filter(cnv_group == "2 + 0")
table(check$nrpcc_bin)

sample_check <- df_merged_sub %>% distinct(sample_id, .keep_all = T)
table(sample_check$nrpcc_bin)

plot_un <- ggplot(df_merged_sub, aes(x = factor(cnv_group), y = hdi_width)) +
  geom_violin(trim = FALSE,  width = 2, fill = "#00A087", alpha = 0.6) +
  geom_boxplot(width = 0.1, outlier.shape = NA, fill = "white", alpha = 0.8) +
  facet_wrap(~ wgd, ncol = 1, ) +
  theme_bw() +
  labs(x = "", y = "Uncertainty", title = "Distribution of Uncertainty (CI = 0.95)") +
  theme(
    axis.text.x = element_text(vjust = 1, hjust = 1, size = 14, colour = "black", angle = 45),
    axis.text.y = element_text(size = 14, colour = "black"),
    axis.title  = element_text(size = 16),
    title = element_text(size = 16),
    plot.margin = margin(10, 10, 10, 10),
    strip.text  = element_text(size = 15),
  ) + 
  scale_y_continuous(breaks = seq(0, 1, by = 0.2), limits = c(0, 1)) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey60")

ggsave(filename = paste0(path_save, "/HDI_uncertainty_wgd05_cnvcom.pdf"),
       plot     = plot_un,
       width = 12, 
       height = 4,
       units    = "in")

## X
plot_un <- ggplot(check, aes(x = factor(nrpcc_bin), y = hdi_width)) +
  geom_violin(trim = FALSE, width = 1.5, fill = "#00A087", alpha = 0.6) +
  geom_boxplot(width = 0.1, outlier.shape = NA, fill = "white", alpha = 0.8) +
  theme_bw() +
  labs(x = "NRPCC", y = "Uncertainty", title = "Distribution of Uncertainty (CI = 0.95)") +
  theme(
    axis.text.x = element_text(vjust = 1, hjust = 1, size = 12, colour = "black", angle = 45),
    axis.text.y = element_text(size = 12, colour = "black"),
    axis.title  = element_text(size = 14),
    title = element_text(size = 16),
    plot.margin = margin(10, 10, 10, 10),
    strip.text  = element_text(size = 15),
  ) + 
  scale_y_continuous(breaks = seq(0, 1, by = 0.2), limits = c(0, 1)) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey60")

ggsave(filename = paste0(path_save, "/HDI_uncertainty_wgd05_cnvcom_nrpcc_x.pdf"),
       plot     = plot_un,
       width = 12, 
       height = 4,
       units    = "in")

```

## NRPCC
```{r}
plot_un <- ggplot(df_merged_NRPCC, aes(x = factor(nrpcc_group), y = hdi_width, fill = factor(nrpcc_group))) +
  geom_violin(trim = FALSE, alpha = 0.6) +
  geom_boxplot(width = 0.2, outlier.shape = NA, fill = "white", alpha = 0.8) +
  scale_y_continuous(breaks = seq(0, 1, by = 0.2), limits = c(0, 1)) +
  scale_fill_manual(values = c("#00A087", "#E64B35", "#4DBBD5", "#F39B7F", "#8491B4"),
                    name = "NRPCC Group") +
  facet_wrap(~ cnv_group, nrow = 1) +
  theme_bw() +
  labs(x = "", y = "Uncertainty", title = "Distribution of Uncertainty (CI = 0.95)") +
  theme(
    axis.text.x  = element_blank(),
    axis.text.y  = element_text(size = 14, colour = "black"),
    axis.title   = element_text(size = 16),
    title        = element_text(size = 16),
    plot.margin  = margin(10, 10, 10, 10),
    strip.text   = element_text(size = 15),
    legend.position = "top",
    legend.text = element_text(size = 12)
  ) +
  geom_hline(yintercept = 0.5, linetype = "dashed", color = "grey60")

ggsave(filename = paste0(path_save, "/HDI_uncertainty_wgd_cnv_nrpcc.pdf"),
       plot     = plot_un,
       width = 20, 
       height = 5,
       units    = "in")
```

# SNV with HDI < 0.5
```{r}
## Subset <0.5
LOH_type <- LOH_type %>% mutate(Tumor_Sample_Barcode = sample_id)

hdi_05 <- hdi_nonsyn %>% filter(hdi_width < 0.5) %>% 
  inner_join(LOH_type, by = "Tumor_Sample_Barcode")

mut_per_sample <- hdi_05 %>%
  group_by(Tumor_Sample_Barcode) %>%
  summarise(
    histology_abbreviation = first(histology_abbreviation),
    n_mutations = n(),
    .groups = "drop"
  ) %>% 
  ungroup() %>%
  mutate(
    histology_abbreviation = factor(
      histology_abbreviation,
      levels = sort(unique(histology_abbreviation))
    )
  )

plot_05 <- ggplot(mut_per_sample,
       aes(x = histology_abbreviation,
           y = n_mutations,
           fill = histology_abbreviation)) +
  geom_boxplot(
    width = 0.7,
    outlier.shape = NA,
    #fill = "grey90",
    color = "black"
  ) +
  geom_jitter(
    width = 0.2,
    height = 0,
    alpha = 0.6,
    size = 0.5
  ) +
  scale_y_log10(
    breaks = c(1, 10, 100, 1000, 10000),
    labels = scales::label_number()
  ) +
  ylab("Number of SNVs per Sample") + geom_hline(
    yintercept = 100,
    linetype = "dashed",
    color = "black",
    linewidth = 0.5
  ) +
  scale_fill_manual(values = type_colors) +
  ggtitle("SNVs with Timing Uncertainty < 0.5") +
  theme_bw() +
  theme(
    axis.text.x = element_text(size = 11, color = "black", hjust = 1, vjust = 1, angle = 45),
    axis.text.y = element_text(size = 11, color = "black"),
    axis.title.x = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5, face = "plain"),
    panel.grid = element_blank(),
    axis.line = element_line(color = "black"),
    axis.ticks = element_line(color = "black"),
    #axis.title.y = element_blank(),
    #axis.ticks.y = element_blank(),
    #axis.line.y = element_blank(),
    legend.position = "none",
    plot.margin = margin(t = 10, b = 10, l = 10, r = 10)
  )

plot_05
path_save = "/Users/wchen20/Desktop/2026-02-16_Draft/Revision/R1Major5/"
ggsave(
  filename = paste0(path_save, "/HDI_uncertainty_05.pdf"),
  plot     = plot_05,
  width = 12, 
  height = 4,
  units    = "in")
```